### 2021-06-10 analysis_eTMB.Rmd
Initial example

---
title: "Analysis of eTMB"
output:
  pdf_document: default
  code_folding: default
  html_document:
    code_folding: hide
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading libraries

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(MutationalPatterns)
library(NMF)
library(reshape2)
library(grid)
library(gridExtra)
```

## Preliminary functions

```{r}
# Loads a mutational profile with columns sample, ext.context and num
# Subsets such file to samples in sample.list and generates a matrix
# with the mutational profile that can be used by MutationalPatterns
load_mutational_matrix<- function(filename, sample.list) {

  mutational.signature <-  read_tsv(filename)
  order.tri <-  unique(mutational.signature$ext.context)
  sample.names <-  unique(mutational.signature$sample)

  mutational.signature <-  mutational.signature %>%
    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
    filter(sample %in% sample.list) %>%
    select(sample,ext.context,num) %>%
    spread(sample,num) %>%
    select(one_of("ext.context",sample.names))

  mut.matrix <- as.matrix(sapply(mutational.signature[,2:ncol(mutational.signature)], as.numeric))  
  rownames(mut.matrix) <-  mutational.signature$ext.context

  # Add 1e-04 for use with NMF
  mut.matrix <-  mut.matrix + 1e-04

  return (mut.matrix)
}

load_old_mutational_matrix<- function(filename,mut.order="supplied_results/mut_sig.order.txt") {

  mutational.signature = read_tsv(filename)
  mut_mat = read_tsv(mut.order)

  mut.join = mutational.signature %>%
    inner_join(mut_mat)

  order.tri = unique(mut.join$ext.context)

  mut.join = mut.join %>%
    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
    select(ext.context,prob)

  mut.matrix=as.matrix(sapply(mut.join[,2:ncol(mut.join)], as.numeric))  
  rownames(mut.matrix) = mut.join$ext.context

  return (t(mut.matrix)[,1:96])
}

# Loads the published COSMIC mutational signatures in a format that can be
# used by MutationalPatterns
load_cosmic_matrix<- function(mut.order="supplied_results/mut_sig.order.txt") {

  sp_url <- paste("https://cancer.sanger.ac.uk/cancergenome/assets/",
                "signatures_probabilities.txt", sep = "")

  cancer_signatures <-  read.table(sp_url, sep = "\t", header = TRUE)
  # Match the order of the mutation types to MutationalPatterns standard
  mut_mat <-  read_tsv(mut.order)
  new_order <-  match(mut_mat$ext.context, cancer_signatures$Somatic.Mutation.Type) 
  # Reorder cancer signatures dataframe
  cancer_signatures <-  cancer_signatures[as.vector(new_order),]
  # Add trinucletiode changes names as row.names
  row.names(cancer_signatures) <-  cancer_signatures$Somatic.Mutation.Type
  # Keep only 96 contributions of the signatures in matrix
  cancer_signatures <-  as.matrix(cancer_signatures[,4:33])
  return(cancer_signatures)
}

# Returns a frame with columns Signature, Sample, Contribution with the absolute
# contribution of each signature for each sample
calc_abs_contribution <- function(nmf.res) {

  total_signatures <-  colSums(nmf.res$signatures) 
  abs_contribution <-  nmf.res$contribution * total_signatures
  rownames(abs_contribution) <- sig.names

  m_contribution <-  melt(abs_contribution)
  colnames(m_contribution) <-  c("Signature", "Sample", "Contribution")

  return (m_contribution)
}

# Returns a frames with the cosine similarity (fit) between the original profile and the
# reconstructed profile
calc_fit_contribution <- function(mut.matrix,nmf.res) {

  # calculate all pairwise cosine similarities
  cos.sim.all <- cos_sim_matrix(mut.matrix, nmf.res$reconstructed)
  # extract cosine similarities per sample between original and reconstructed
  fit.table<- as.data.frame(diag(cos_sim_ori_rec))
  fit.table$Sample <-  rownames(fit.table)
  colnames(fit.table) <-  c("fit","sample")

  return(fit.table)
}

# Returns with the total mutations per case after removing the FFPE component
remove_ffpe_contribution <- function(contribution) {

  norm.contribution <-  contribution %>%
    filter(Signature != "FFPE") %>%
    group_by(Sample) %>%
    summarize(norm.mut=sum(Contribution))

  return (norm.contribution)
}

calc_mmr_contribution <- function(contribution) {

  norm.contribution <-  contribution %>%
    filter(Signature == "MMR") %>%
    group_by(Sample) %>%
    summarize(mmr.mut=sum(Contribution))

  return (norm.contribution)
}

# Summarize cohort according to mmr.status, sex, tissue and tissue of origin
summarize_cohort <-function(mmr.table) {

  summary.age <-  mmr.table %>%
    group_by(mmr.type) %>%
      summarize(n=n(),
            avg.age=mean(Age.DX,na.rm=TRUE),std.age=sd(Age.DX,na.rm=TRUE),
            min.age=min(Age.DX,na.rm=TRUE),max.age=max(Age.DX,na.rm=TRUE))

  summary.sex <- 
    mmr.table %>%
      group_by(mmr.type,Sex) %>%
      summarize (n=n()) %>%
      spread (Sex,n)

  summary.tissue <-  
    mmr.table %>%
    group_by(mmr.type,Tissue) %>%
    summarize (n=n()) %>%
    spread(Tissue,n)

  summary <-  summary.age %>%
    inner_join(summary.sex) %>%
    inner_join(summary.tissue)

  return(summary)
}

```

## Sample filtering

We exclude samples that meet the following criteria:

 * Their sequencing had less than 10 million reads.
 * Samples from our pilot with incomplete information.
 * Have no corresponding normal tissue (only tumor).
 * Samples with unknown tissue type.
 * Samples with wrong MSI/MSS phenotype. 
 * Samples that did not meet age criteria.
 * Samples from ovarian cancer or small bowel cancer.

```{r message=FALSE}
mapping.stats <- read_tsv("supplied_results/master_statistics_table.txt") %>%
  select(sample.name,tumor.status,stat.Total_Reads)

samples.pilot <-  c("MSI_MLH1p_1")
samples.na.tissue <-  c("MSI_MLH1G_2","MSI_MLH1G_6","MSI_MSH2_11","MSS_1")
samples.exclude.age <-  c("MSI_MLH1HM_14")
samples.wrong.phenotype <-  c("MSI_MLH1G_9")
samples.small.bowel <- c("MSI_PMS2_12","MSS_9")

min.reads <- 1e7 #minimum number of reads

mapping.info.filter <-  mapping.stats %>%
  spread(tumor.status,stat.Total_Reads) %>%
  filter(!is.na(normal)) %>%
  filter(normal>min.reads & tumor>min.reads) %>%
  filter(!sample.name %in% samples.pilot) %>%
  filter(!sample.name %in% samples.na.tissue) %>%
  filter(!sample.name %in% samples.exclude.age) %>%
  filter(!sample.name %in% samples.wrong.phenotype) %>%
  filter(!sample.name %in% samples.small.bowel)

```

## Initial NMF analysis

We start the analysis by loading the mutational profiles from the samples that made it into `mapping.info.filter`

```{r warning=FALSE, message=FALSE}
mut.matrix <-  load_mutational_matrix("supplied_results/signatures.txt",
  mapping.info.filter$sample.name)
```

And we apply non-negative matrix factorization

```{r}
nmf.res <- extract_signatures(mut.matrix, rank = 4, nrun = 10)
```

And plot the 4 signatures that we found in the processes side

```{r}
plot_96_profile(nmf.res$signatures, condensed = TRUE)
```

To infer what process corresponds to what signature, we load the COSMIC v2 mutational signatures and compare them against our signatures

```{r message=FALSE}
# Plot COSMIC v2 according to their similarity to each other
cosmic.signatures <- load_cosmic_matrix()
hclust_cosmic <-  cluster_signatures(cosmic.signatures,
                                   method = "average")
cosmic_order <-  colnames(cosmic.signatures)[hclust_cosmic$order]
plot(hclust_cosmic)

# Do a heatmap showing similarity between COSMIC v2 and our signatures
colnames(nmf.res$signatures) <- seq (1:4)
cos_sim_samples_signatures <-  cos_sim_matrix(nmf.res$signatures, cosmic.signatures)
plot_cosine_heatmap(cos_sim_samples_signatures,
                    col_order = cosmic_order,
                    cluster_rows = FALSE)
```

From the figure we get that:

 * Our first signature corresponds to COSMIC signature 10 with a cosine similarity of `r round(cos_sim(cosmic.signatures[,10],nmf.res$signatures[,1]),2)`

 * The second signature corresponds to COSMIC signature 5 with a cosine similarity of `r round(cos_sim(cosmic.signatures[,5],nmf.res$signatures[,2]),2)`

 * The third signature has a high similarity with COSMIC signature 30 with a cosine similarity of `r round(cos_sim(cosmic.signatures[,30],nmf.res$signatures[,3]),2)`. We will come to that later on.

* The fourth signature corresponds to COSMIC signature 6 with a cosine similarity of `r round(cos_sim(cosmic.signatures[,6],nmf.res$signatures[,4]),2)`

On further inspection we compare our third signature against a previously identified FFPE signature

```{r warning=FALSE, message=FALSE}
ffpe.aditya <-  load_old_mutational_matrix("supplied_results/signatures/ffpe.aditya.txt")
```

And we obtain that it has a cosine similarity of `r round(cos_sim(ffpe.aditya,nmf.res$signatures[,3]),2)` against this signature.

Given this information we label each of our resulting signatures:

```{r}
sig.names <-  c("PolE","Aging","FFPE","MMR")
colnames(nmf.res$signatures) <- sig.names
rownames(nmf.res$contribution) <- sig.names

sig.order <-  c("MMR","Aging","PolE","FFPE")
inferred.signatures =nmf.res$signatures[,sig.order]
```

## Plotting samples with high TMB

It seems samples with unusually high TMB have similar mutational profiles. Let's plot the top-6 highest mutated samples in a single graph

```{r message=FALSE, warning=FALSE}
# Get tmb for each sample that was used for mutational analysis
burden.table <-  read_tsv("supplied_results/burden.final.txt") %>%
    mutate(burden = (somatic_snvs/tumor_doc_50X) * 1e6) %>%
    filter(sample %in% colnames(nmf.res$contribution))

# Format sample name to remove MSI prefix
sample.names <-  colnames(nmf.res$contribution) %>%
  str_replace('MSI_','') %>%
  str_replace('MSH6p','MSH6')

# Table with tmb info and short name as a factor sorted by tmb (desc)
burden.table <-  burden.table %>%
    mutate(short.sample = str_replace(sample,"MSI_","")) %>%
    mutate(short.sample = str_replace(short.sample,"MSH6p","MSH6")) %>%
    mutate(short.sample = factor(short.sample,levels=rev(sample.names))) %>%
    arrange(desc(burden))

# Samples sorted by tmb in descending order
pch1 <- burden.table %>%
  ggplot(aes(x=reorder(short.sample,burden),y=burden,fill=type))+
  geom_bar(stat="identity")+
  coord_flip()+
  ylab("Tumor mutational burden")+
  xlab("Sample")

high.burden.samples <-  burden.table %>%
  head(n=6) %>%
  select(sample)

mut.high = load_mutational_matrix("supplied_results/signatures.txt",
  high.burden.samples$sample)

pch2 <- plot_96_profile(mut.high[], condensed = TRUE)

supp.1.grob = grid.arrange(pch1,pch2, ncol=2)

supp.1.grob
#ggsave(file="figures/supp.1.png",
       #supp.1.grob, width=12, height=7, units="in", dpi=300)

```

## Plotting the NMF analysis

```{r}
pch1 <- plot_96_profile(inferred.signatures, condensed = TRUE)

# Get absolute contribution for each signature x sample
sig.abs.contribution = calc_abs_contribution (nmf.res)

# Calculate absolute tmb and pct tmb for each sample x signature 
sample.sig.table = sig.abs.contribution %>%
    inner_join(burden.table, by=c("Sample"="sample")) %>%
    mutate(burden.contribution = Contribution/tumor_doc_50X * 1e6,
           burden.pct = burden.contribution/burden,
           Signature = factor(Signature,levels=sig.order))

# Plot absolute tmb contribution for each sample x signature in waterfall diagram
pch2 <- ggplot(sample.sig.table,
               aes (x=reorder(short.sample,burden),
                    y=burden.contribution,fill=Signature))+
  geom_bar(stat="identity")+
  coord_flip()+
  ylab("Tumor mutational burden")+
  xlab("Sample")+
  theme(legend.position="none")

fig.1.grob = grid.arrange(pch1,pch2, ncol=2)
fig.1.grob
```