### 2021-06-17 Function.Rmd

---
title: "Function"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(MutationalPatterns)
library(BSgenome)
library(BiocManager)
install("BSgenome.Hsapiens.UCSC.hg19")
ref_genome <- "BSgenome.Hsapiens.UCSC.hg19"
library(ref_genome, character.only = TRUE)
```

```{r}
vcf_files1 <- list.files(system.file("extdata", package="MutationalPatterns"), pattern = ".vcf", full.names = TRUE)
#vcf_files <- vcf_files1[-c(1,2,3,7)]
vcf_files <-vcf_files1[c(4,5,6)]
sample_names <- c("colon1", "colon2", "colon3")
vcfs <- read_vcfs_as_granges(vcf_files, sample_names, ref_genome)
tissue <- c(rep("colon",3))
```

```{r}
mut_mat <- mut_matrix(vcf_list = vcfs, ref_genome = ref_genome)
mut_mat <- mut_mat + 0.0001
```

```{r}
nmf_res <- extract_signatures(mut_mat, rank = 2, nrun = 10)
colnames(nmf_res$signatures) <- c("Signature A", "Signature B")
rownames(nmf_res$contribution) <- c("Signature A", "Signature B")
```

```{r}
#Goal: Pr(Signature|Mutation) in a particular sample
#Change name for nmf_matrix
#Split denominator into two pieces, add a parameter for which method we use
#One function that doesn't need mutational signature
Bayes_function <- function(mutation, signature, sample, nmf_matrix, mutation_matrix){
  #Pr(Mutation|Signature)
  a <- nmf_matrix$signatures[mutation,signature]/sum(nmf_matrix$signatures[,signature])
  #Pr(Signature)
  b <- nmf_matrix$contribution[signature,sample]/sum(nmf_matrix$contribution[,sample])
  #Pr(Mutation)
  c <- mutation_matrix[mutation, sample]/sum(mutation_matrix[,sample])
  #Bayes Formula Calculation
  prob <- a*b/c
  return(prob)
}
```

```{r}
Bayes_function <- function(mutation, signature, sample, nmf_matrix, mutation_matrix){
  #Pr(Mutation|Signature)
  prob_mut_given_sig <- nmf_matrix$signatures[mutation,signature]/sum(nmf_matrix$signatures[,signature])
  #Pr(Signature)
  prob_sig <- nmf_matrix$contribution[signature,sample]/sum(nmf_matrix$contribution[,sample])
  #Pr(Mutation)

  #Bayes Formula Calculation
  prob <- a*b/c
  return(prob)
}
```

```{r}
#Goal: Pr(SignatureA|A[C>A]G) and Pr(SignatureB|A[C>A]G) for sample colon1
Bayes_function("A[C>A]G", "Signature A", "colon1", nmf_res, mut_mat)
Bayes_function("A[C>A]G", "Signature B", "colon1", nmf_res, mut_mat)
```

```{r}
plot_96_profile(nmf_res$signatures)
```

```{r}
Bayes_function("G[C>T]G", "Signature A", "colon1", nmf_res, mut_mat)
Bayes_function("G[C>T]G", "Signature B", "colon1", nmf_res, mut_mat)
```