### 2021-06-17 Bayes Function Testing.Rmd
testing function on samples created from extreme proportions of two signatures

---
title: "Bayes Function Testing"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MutationalPatterns)
library(BSgenome)
library(BiocManager)
ref_genome <- "BSgenome.Hsapiens.UCSC.hg19"
library(ref_genome, character.only = TRUE)
source("bayes_function.R")
```

```{r}
# Loads the published COSMIC mutational signatures in a format that can be
# used by MutationalPatterns
load_cosmic_matrix<- function(mut.order="supplied_results/mut_sig.order.txt") {

  sp_url <- paste("https://cancer.sanger.ac.uk/cancergenome/assets/",
                "signatures_probabilities.txt", sep = "")

  cancer_signatures <-  read.table(sp_url, sep = "\t", header = TRUE)
  # Match the order of the mutation types to MutationalPatterns standard
  mut_mat <-  read_tsv(mut.order)
  new_order <-  match(mut_mat$ext.context, cancer_signatures$Somatic.Mutation.Type) 
  # Reorder cancer signatures dataframe
  cancer_signatures <-  cancer_signatures[as.vector(new_order),]
  # Add trinucletiode changes names as row.names
  row.names(cancer_signatures) <-  cancer_signatures$Somatic.Mutation.Type
  # Keep only 96 contributions of the signatures in matrix
  cancer_signatures <-  as.matrix(cancer_signatures[,4:33])
  return(cancer_signatures)
}
```

```{r}
create_test_sample <- function(signatures, prop, num) {
  cosmic_signatures <- load_cosmic_matrix()
  process <- cosmic_signatures[, signatures]

  exposures <- matrix(prop, nrow = length(signatures), ncol = num)

  return ((process %*% exposures) * 10^5)
}
```

```{r}
#Create a set of 4 test samples with extreme propotions of COSMIC signatures 18 and 21
test_sample <- create_test_sample(c(18,21), c(1,0,0,1,.9,.1,.1,.9),4)
test_sample
```

```{r}
#Run NMF analysis on signatures
test_nmf <- extract_signatures(test_sample, rank = 2, nrun = 10)

#Plot signatures to match with signatures 18 and 21: V1 appears to match signature 21, V2 appears to match signature 18
plot_96_profile(test_nmf$signatures)

#set column and row names to names of signatures
colnames(test_nmf$signatures) <- c("Signature 21", "Signature 18")
rownames(test_nmf$contribution) <- c("Signature 21", "Signature 18")
```

```{r}
test_nmf$contribution
plot_contribution(test_nmf$contribution, test_nmf$signature, mode = "relative", coord_flip = TRUE)
```

```{r}
#G[C>A]A mutation is the most common in Signature 18
#Sample 1 is ~100% Signature 18
extract_all_prob("G[C>A]A", 1, test_nmf)
#Sample 3 is ~90% Signature 18
extract_all_prob("G[C>A]A", 3, test_nmf)
```

```{r}
#G[T>C]A mutation is the most common in Signature 21
#Sample 2 is ~100% Signature 21
extract_all_prob("G[T>C]A", 2, test_nmf)
#Sample 4 is ~90% Signature 21
extract_all_prob("G[T>C]A", 4, test_nmf)
```