### 2021-06-18 Bayes Function Testing.Rmd
Updated testing with finer range of exposures and expected output for each chunk.

---
title: "Bayes Function Testing"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MutationalPatterns)
library(BSgenome)
library(BiocManager)
ref_genome <- "BSgenome.Hsapiens.UCSC.hg19"
library(ref_genome, character.only = TRUE)
source("bayes_function.R")
```

```{r}
# Loads the published COSMIC mutational signatures in a format that can be
# used by MutationalPatterns
load_cosmic_matrix<- function(mut.order="supplied_results/mut_sig.order.txt") {

  sp_url <- paste("https://cancer.sanger.ac.uk/cancergenome/assets/",
                "signatures_probabilities.txt", sep = "")

  cancer_signatures <-  read.table(sp_url, sep = "\t", header = TRUE)
  # Match the order of the mutation types to MutationalPatterns standard
  mut_mat <-  read_tsv(mut.order)
  new_order <-  match(mut_mat$ext.context, cancer_signatures$Somatic.Mutation.Type) 
  # Reorder cancer signatures dataframe
  cancer_signatures <-  cancer_signatures[as.vector(new_order),]
  # Add trinucletiode changes names as row.names
  row.names(cancer_signatures) <-  cancer_signatures$Somatic.Mutation.Type
  # Keep only 96 contributions of the signatures in matrix
  cancer_signatures <-  as.matrix(cancer_signatures[,4:33])
  return(cancer_signatures)
}
```

```{r}
#Creates samples/mutational profiles using selected signatures from COSMIC
#and desired contributions from each signature.
create_test_sample <- function(signatures, prop, num) {
  cosmic_signatures <- load_cosmic_matrix()
  process <- cosmic_signatures[, signatures]

  exposures <- matrix(prop, nrow = length(signatures), ncol = num)

  return ((process %*% exposures) * 10^5)
}
```

```{r}
#Create a set of 7 test samples with extreme proportions of COSMIC signatures 18 and 21
test_sample <- create_test_sample(c(18,21), c(  1,  0,
                                              0.9,0.1,
                                              0.7,0.3,
                                              0.5,0.5,
                                              0.3,0.7,
                                              0.1,0.9,
                                                0,  1),
                                  num = 7)
test_sample
```

```{r}
#Run NMF analysis on signatures
test_nmf <- extract_signatures(test_sample, rank = 2, nrun = 10)

#Plot signatures to match with signatures 18 and 21: V1 appears to match signature 21, V2 appears to match signature 18
plot_96_profile(test_nmf$signatures)

#set column and row names to names of signatures
colnames(test_nmf$signatures) <- c("Signature 18", "Signature 21")
rownames(test_nmf$contribution) <- c("Signature 18", "Signature 21")
```

```{r}
test_nmf$contribution
plot_contribution(test_nmf$contribution, test_nmf$signature, mode = "relative", coord_flip = TRUE)
```

## Proportions should lean in favor for signature 18 and decrease with its sample size

```{r}
#G[C>A]A mutation is the most common in Signature 18
#1287 times more likely in signature 18 than 21
#Sample 1 is ~100% Signature 18
extract_all_prob("G[C>A]A", 1, test_nmf)
#Sample 2 is ~90% Signature 18
extract_all_prob("G[C>A]A", 2, test_nmf)
#Sample 3 is ~70% Signature 18
extract_all_prob("G[C>A]A", 3, test_nmf)
#Sample 3 is ~50% Signature 18
extract_all_prob("G[C>A]A", 3, test_nmf)
#Sample 4 is ~30% Signature 18
extract_all_prob("G[C>A]A", 4, test_nmf)
#Sample 5 is ~10% Signature 18
extract_all_prob("G[C>A]A", 5, test_nmf)
#Sample 6 is ~0% Signature 18
extract_all_prob("G[C>A]A", 6, test_nmf)
```

## Proportions should lean in favor for signature 21 and decrease with its sample size

```{r}
#G[T>C]A mutation is the most common in Signature 21
#269 times more likely in signature 21 than 18
#Sample 6 is ~100% Signature 21
extract_all_prob("G[T>C]A", 6, test_nmf)
#Sample 5 is ~90% Signature 21
extract_all_prob("G[T>C]A", 5, test_nmf)
#Sample 4 is ~70% Signature 21
extract_all_prob("G[T>C]A", 4, test_nmf)
#Sample 4 is ~50% Signature 21
extract_all_prob("G[T>C]A", 4, test_nmf)
#Sample 3 is ~30% Signature 21
extract_all_prob("G[T>C]A", 3, test_nmf)
#Sample 2 is ~10% Signature 21
extract_all_prob("G[T>C]A", 2, test_nmf)
#Sample 1 is ~0% Signature 21
extract_all_prob("G[T>C]A", 1, test_nmf)
```

```{r}
#Find smallest difference between contribution rate for every mutation type between
#signature 18 and 21. Attempting to find mutation type w/ equal probabilities for both.
cosmic_signatures <- load_cosmic_matrix()
(process18 <- cosmic_signatures[, c(18)])
(process21 <- cosmic_signatures[, c(21)])
diff_signatures <- abs(process21-process18)
diff_signatures[93]=1 #contribution for signature 21 is 0 (T[T>G]A)
diff_signatures[85]=1 #contribution for signature 21 is 0
which.min(diff_signatures)
```

## Signature 21 probabilities should be zero for all samples

```{r}
#LOW probability of T[T>G]A mutation in signature 18
#zero probability of T[T>G]A mutation in signature 21
#100% Signature 18 and 0% Signature 21
extract_all_prob("T[T>G]A", 1, test_nmf)
#90% Signature 18 and 10% Signature 21
extract_all_prob("T[T>G]A", 2, test_nmf)
#70% Signature 18 and 30% Signature 21
extract_all_prob("T[T>G]A", 3, test_nmf)
#50% Signature 18 and 50% Signature 21
extract_all_prob("T[T>G]A", 4, test_nmf)
#30% Signature 18 and 70% Signature 21
extract_all_prob("T[T>G]A", 5, test_nmf)
#10% Signature 18 and 90% Signature 21
extract_all_prob("T[T>G]A", 6, test_nmf)
#0% Signature 18 and 100% Signature 21
extract_all_prob("T[T>G]A", 7, test_nmf) #BAD RESULT
```

## Probabilities should almost reflect contribution proportions

```{r}
#Low chances of C[C>G]A mutations in both but roughly equal chances of mutations
#(4.58 more likely in signature 18, as opposed to hundreds or thousands times more likely)
#100% Signature 18 and 0% Signature 21
extract_all_prob("C[C>G]A", 1, test_nmf)
#90% Signature 18 and 10% Signature 21
extract_all_prob("C[C>G]A", 2, test_nmf)
#70% Signature 18 and 30% Signature 21
extract_all_prob("C[C>G]A", 3, test_nmf)
#50% Signature 18 and 50% Signature 21
extract_all_prob("C[C>G]A", 4, test_nmf)
#30% Signature 18 and 70% Signature 21
extract_all_prob("C[C>G]A", 5, test_nmf)
#10% Signature 18 and 90% Signature 21
extract_all_prob("C[C>G]A", 6, test_nmf)
#0% Signature 18 and 100% Signature 21
extract_all_prob("C[C>G]A", 7, test_nmf)
```