### 2021-06-21 Sample Distribution Testing.Rmd
resolved function errors and tested on FFPE and PolE signature distributions

---
title: "Sample Distribution Testing"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(MutationalPatterns)
library(NMF)
source("simulation.R")
```

```{r}
load_old_mutational_matrix<- function(filename,mut.order="supplied_results/mut_sig.order.txt") {

  mutational.signature = read_tsv(filename)
  mut_mat = read_tsv(mut.order)

  mut.join = mutational.signature %>%
    inner_join(mut_mat)

  order.tri = unique(mut.join$ext.context)

  mut.join = mut.join %>%
    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
    select(ext.context,prob)

  mut.matrix=as.matrix(sapply(mut.join[,2:ncol(mut.join)], as.numeric))  
  rownames(mut.matrix) = mut.join$ext.context

  return (t(mut.matrix)[,1:96])
}

#load ffpe signature
ffpe.signature <-  load_old_mutational_matrix("supplied_results/ffpe.signature.txt")
ffpe.signature
```

```{r}
#convert ffpe.signature to a matrix to be compatible with MutationalPatterns
ffpe.signature.matrix <- as.matrix(ffpe.signature)
plot_96_profile(ffpe.signature.matrix)
```

```{r}
#create ffpe sample vector of 1000 mutations
ffpe.sample.vector <- create_signature_sample_vector(ffpe.signature.matrix, 1000)
ffpe.sample.vector
```

```{r}
#create sample matrix to match ffpe signature
ffpe.sample.matrix <- create_signature_sample_matrix(ffpe.signature.matrix, 10000)
ffpe.sample.matrix
#calculate the cosine similarity between sample matrix and original ffpe matrix
cos_sim_matrix(ffpe.sample.matrix, ffpe.signature.matrix)
#plot original ffpe matrix
plot_96_profile(ffpe.signature.matrix)
#plot our ffpe sample matrix
plot_96_profile(ffpe.sample.matrix)
```

```{r}
#load COSMIC signatures
cosmic.sigs <- load_cosmic_matrix()
```

```{r}
polE <- as.matrix(cosmic.sigs[,10])
polE
plot_96_profile(polE)

polE.sample.vector <- create_signature_sample_vector(polE, 1000)
polE.sample.vector

polE.sample.matrix <- create_signature_sample_matrix(polE, 1000)
polE.sample.matrix
cos_sim_matrix(polE, polE.sample.matrix)
plot_96_profile(polE.sample.matrix)
```

```