### 2021-06-23 manual_classification.Rmd
Progressed in classification of signature in simulated mutations

---
title: "multiple_sig_sample"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(MutationalPatterns)
library(NMF)
library(tidyverse)

source("simulation.R")
source("bayes_function.R")
```

```{r}
cosmic.sigs <- load_cosmic_matrix()
ffpe.sig <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))
```

```{r}
cosmic.sig6 <- as.matrix(cosmic.sigs[, 6])
cosmic.sig6

sample.sig6 <- create_signature_sample_vector(cosmic.sig6, 100)
sig6.truth <- rep("MMR", 100)
sample.ffpe <- create_signature_sample_vector(ffpe.sig, 100)
ffpe.truth <- rep("FFPE", 100)
```

```{r}
# Create ground truth column

df.final <- data.frame(append(sample.sig6, sample.ffpe), append(sig6.truth, ffpe.truth))
df.final
```

```{r}
signature.sample <- append(sample.sig6, sample.ffpe)

#converts sample vector into a table of probabilities
tab <- table(signature.sample)/sum(table(signature.sample))
tab
#converts table of probabilities to a data frame and defines column names
sample.probs.df <- data.frame(tab)
colnames(sample.probs.df) <- c("mutations", "frequencies")
sample.probs.df

#defines a function that joins two data frames and replaces NA with 0
left_join_NA <- function(x, y, by) {
  left_join(x = x, y = y, by = by) %>% 
    mutate_each(funs(replace(., which(is.na(.)), 0)))
}

#creates a vector of mutation types
mutations <- rownames(cosmic.sig6)

#creates a data.frame of mutation types to join with sample probabilities
df <- data.frame(mutations)
df

#joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
sample.df <- left_join_NA(df, sample.probs.df, by = "mutations")
sample.df

#creates a matrix of sample frequencies and defines row names as mutation types
sample.matrix<- as.matrix(sample.df$frequencies) + 0.0001
rownames(sample.matrix)= mutations

plot_96_profile(sample.matrix)

nmf_res <- fit_to_signatures(sample.matrix, cbind(cosmic.sig6, ffpe.sig))
nmf_res$signatures <- cbind(cosmic.sig6, ffpe.sig)
nmf_res
```

```{r}
colnames(nmf_res$signatures) <- c("MMR", "FFPE")
rownames(nmf_res$contribution) <- c("MMR", "FFPE")

plot_96_profile(nmf_res$signatures)

mmr.prob <- vector(mode = "double")
ffpe.prob <- vector(mode = "double")

for (x in mutations) {
  probs <- extract_all_prob(x, 1, nmf_res)
  print(x)
  print(probs)
  mmr.prob <- c(mmr.prob, probs$MMR)
  ffpe.prob <- c(ffpe.prob, probs$FFPE)
}

mmr.prob
ffpe.prob
```