### 2021-06-29 Varying Signature Function Testing.Rmd
plotted average misclassification rate vs. cosine similarity of COSMIC signatures to FFPE, implemented a function to plot an ROC curve for a simulated sample containing mutations from FFPE and one other COSMIC signature

---
title: "Untitled"
author: "Audrey Mitchell"
date: "6/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MutationalPatterns)
library(NMF)
library(tidyverse)
library(ROCR)
source("bayes_function.R")
source("get_classification_df.R")
```

```{r}
#Load FFPE signature
##Function to load FFPE signature
load_old_mutational_matrix<- function(filename,mut.order="supplied_results/mut_sig.order.txt") {

  mutational.signature = read_tsv(filename)
  mut_mat = read_tsv(mut.order)

  mut.join = mutational.signature %>%
    inner_join(mut_mat)

  order.tri = unique(mut.join$ext.context)

  mut.join = mut.join %>%
    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
    select(ext.context,prob)

  mut.matrix=as.matrix(sapply(mut.join[,2:ncol(mut.join)], as.numeric))  
  rownames(mut.matrix) = mut.join$ext.context

  return (t(mut.matrix)[,1:96])
}

ffpe.signature <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))
```

```{r}
#create vector of the 96 mutation types
mutations <- rownames(ffpe.signature) 
```

```{r}
#Load COSMIC signatures (version 3)
cosmic.signatures <- get_known_signatures()
```

```{r}
#Create vector of the signature names
signature.names <- colnames(cosmic.signatures)
```

```{r}
#Create list of signature matrices with assigned rownames
signature.matrix.list <- list()
for (i in 1:60){
  sig <- cosmic.signatures[,i]
  sig.mat <- as.matrix(sig)
  rownames(sig.mat) <- mutations
  colnames(sig.mat) <- signature.names[i]
  signature.matrix.list[[i]] <- sig.mat
}
```

```{r}
#outer for loop: loop through each COSMIC v3 signature
average.misclassification.rates <- c()
for (x in 1:60){
  signature.misclass.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 180 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],180)
    #Create FFPE sample vector of 20 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 20)
    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    misclassification.rate <- mean(classification.df$misclassification)
    signature.misclass.vec[y] <- misclassification.rate
  }
  average.signature.misclass <- mean(signature.misclass.vec)
  average.misclassification.rates[x] <- average.signature.misclass
}
```

```{r}
#create vector of cosine similarities to the FFPE signature
ffpe.cosmic.cos.sims <- c()
for (i in 1:60){
  cos.sim <- cos_sim_matrix(signature.matrix.list[[i]], ffpe.signature)
  ffpe.cosmic.cos.sims[i] <- cos.sim[1,1]
}
```

```{r}
cos.sim.misclass.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.misclassification.rates)
colnames(cos.sim.misclass.data) <- c("signature", "cosine_similarity", "average_misclassification")
```

```{r}
ggplot(cos.sim.misclass.data, aes(x = cosine_similarity, y = average_misclassification))+
  geom_point()+
  geom_smooth()+
  labs(title = "Average misclassification rate vs. cosine similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average misclassification rate (10 iterations)")
```

```{r}
model <- lm(average_misclassification ~ log(cosine_similarity), data = cos.sim.misclass.data)
summary(model)
```

#ROC Curves

```{r}
df <- data.frame(signature.names, ffpe.cosmic.cos.sims)
colnames(df) <- c("signature", "cos_similarity") 
df <- df %>%
  filter(signature =="SBS17b" | signature == "SBS12" | signature == "SBS41" | signature == "SBS44" | signature == "SBS30")
df
```

```{r}
#ROC curve function
##function(COSMIC signature, number of COSMIC mutations, number of FFPE mutations, number of iterations)
plot_bayes_roc <- function(cosmic_sig_matrix, cosmic_sig_name, cosmic_mutation_count, ffpe_mutation_count, iterations){
  true.pos.vector <- c()
  false.pos.vector <- c()
  for (i in 1:iterations){
    rownames(cosmic_sig_matrix) <- mutations
    sample <-create_signature_sample_vector(cosmic_sig_matrix, cosmic_mutation_count)
    ffpe.sample <- create_signature_sample_vector(ffpe.signature, ffpe_mutation_count)
    class.df <- get_classification_df(list(sample, ffpe.sample), c(cosmic_sig_name, "FFPE"), list(cosmic_sig_matrix, ffpe.signature))
    true.pos.df <- class.df %>%
    filter(classify == cosmic_sig_name & truth == cosmic_sig_name)
  true.pos <- nrow(true.pos.df)/nrow(class.df)
  true.pos.vector[i] <- true.pos
  false.pos.df <- class.df %>%
    filter(classify == cosmic_sig_name & truth == "FFPE")
  false.pos <- nrow(false.pos.df)/nrow(class.df)
  false.pos.vector[i] <- false.pos
  }
  plot(false.pos.vector, true.pos.vector, xlab = "False Positivity Rate", ylab = "True Positivity Rate")
}

```
#MMR (Signature 6, cos similarity to ffpe = )
```{r}
plot_bayes_roc(cosmic.6, "MMR", 180, 20, 200)
```

##COSMIC signature SBS17b (minimum cosine similarity to FFPE signature: 0.00206)

```{r}
cosmic.17b <- as.matrix(cosmic.signatures[,"SBS17b"])
plot_bayes_roc(cosmic.17b, "SBS17b", 180, 20, 200)
```

##COSMIC signature SBS12 (~25th percentile for cosine similarity to FFPE: 0.05969503)

```{r}
cosmic.12 <- as.matrix(cosmic.signatures[,"SBS12"])
plot_bayes_roc(cosmic.12, "SBS12", 180, 20, 200)
```

##COSMIC signature SBS 41 (~median cosine similarity to FFPE signature: 0.148913)

```{r}
cosmic.41 <- as.matrix(cosmic.signatures[,"SBS41"])
plot_bayes_roc(cosmic.41, "SBS41", 180, 20, 200)
```

##COSMIC signature 44 (~75th percentile cosine similarity to FFPE signature: 0.486769)

```{r}
cosmic.44 <- as.matrix(cosmic.signatures[,"SBS44"])
plot_bayes_roc(cosmic.44, "SBS44", 180, 20, 200)
```

##COSMIC signature 30 (maximum cosine similarity to FFPE signature: 0.905587)

```{r}
cosmic.30 <- as.matrix(cosmic.signatures[,"SBS30"])
plot_bayes_roc(cosmic.30, "SBS30", 180, 20, 200)
```