### 2021-07-07 plots.Rmd
functioning shiny app, new function to remove intermediate steps in the shiny app

---
title: "Plots"
author: "Audrey Mitchell"
date: "6/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(NMF)
library(tidyverse)
library(ROCR)
source("bayes_function.R")
source("get_classification_df.R")
```

```{r}
#load FFPE signature as a matrix

##Function to load FFPE signature
load_old_mutational_matrix<- function(filename,mut.order="supplied_results/mut_sig.order.txt") {

  mutational.signature = read_tsv(filename)
  mut_mat = read_tsv(mut.order)

  mut.join = mutational.signature %>%
    inner_join(mut_mat)

  order.tri = unique(mut.join$ext.context)

  mut.join = mut.join %>%
    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
    select(ext.context,prob)

  mut.matrix=as.matrix(sapply(mut.join[,2:ncol(mut.join)], as.numeric))  
  rownames(mut.matrix) = mut.join$ext.context

  return (t(mut.matrix)[,1:96])
}

ffpe.signature <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))
```

```{r}
#create vector of the 96 mutation types
mutations <- rownames(ffpe.signature) 
```

```{r}
#Load COSMIC signatures (version 3)
cosmic.signatures <- get_known_signatures()

#get MMR signature (SBS6)
cosmic.6 <- as.matrix(cosmic.signatures[,"SBS6"])
rownames(cosmic.6) <- mutations
```

```{r}
mmr.sample <-create_signature_sample_vector(cosmic.6, 50)

ffpe.sample <- create_signature_sample_vector(ffpe.signature, 50)

class.df <- get_classification_df(list(mmr.sample, ffpe.sample), c("MMR", "FFPE"), list(cosmic.6, ffpe.signature))
```

```{r}
class.df2 <- class.df %>%
  mutate(mmr.indicator = ifelse(truth == "MMR", 1, 0))
```

```{r}
predictions <- class.df2$MMR
labels <- class.df2$mmr.indicator
pred <- prediction(predictions, labels)
pred
```

```{r}
perf <- performance(pred, "tpr", "fpr")
perf
```

```{r}
plot(perf,
     avg="threshold")
```

```{r}
auc.perf = performance(pred, measure = "auc")
auc.perf@y.values[[1]][1]
```

#Plotting Reconstructed Profiles
- Create a mutational profile without FFPE -> plot mutational profile
- Add FFPE -> plot mutational profile
- Run get_classification_df() function on the FFPE-added mutational profile
- Filter out all the mutations that our function classifies as FFPE (Pr(FFPE) > .5)
- Calculate cosine similarity between original mutational profile and reconstructed

##Create a mutational profile without FFPE
```{r}
#load COSMIC v3 
cosmic.v3 <- get_known_signatures(muttype = "snv")

#create POLE signature matrix
pole.signature <- as.matrix(cosmic.v3[,"SBS10b"])
rownames(pole.signature) <- mutations

#create POLE sample vector
pole.sample.vector <- create_signature_sample_vector(pole.signature, 100)

#create MMR signature matrix
mmr.signature <- as.matrix(cosmic.v3[,"SBS6"])
rownames(mmr.signature) <- mutations

#create MMR sample vector
mmr.sample.vector <- create_signature_sample_vector(mmr.signature, 100)

#Combine POLE and MMR vector
sample.vector <- c(pole.sample.vector, mmr.sample.vector)

#transform sample.vector into a matrix of probabilities
tab <- table(sample.vector)/sum(table(sample.vector))
##converts table of probabilities to a data frame and defines column names
sample.probs.df <- data.frame(tab)
colnames(sample.probs.df) <- c("mutations", "frequencies")

#defines a function that joins two data frames and replaces NA with 0
left_join_NA <- function(x, y, by) {
  left_join(x = x, y = y, by = by) %>% 
  mutate_each(funs(replace(., which(is.na(.)), 0)))
}

#creates a data.frame of mutation types to join with sample probabilities
df <- data.frame(mutations)

#joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
sample.df <- left_join_NA(df, sample.probs.df, by = "mutations")

#Add pseudocount to column of probabilities 
pole.mmr.matrix<- as.matrix(sample.df$frequencies) + 0.0001
rownames(pole.mmr.matrix)= mutations

plot_96_profile(pole.mmr.matrix)
```

#Add FFPE to the sample
```{r}
ffpe.sample.vector <- create_signature_sample_vector(ffpe.signature, 50)

new.sample.vector <- c(sample.vector, ffpe.sample.vector)

#transform sample.vector into a matrix of probabilities
tab <- table(new.sample.vector)/sum(table(new.sample.vector))
##converts table of probabilities to a data frame and defines column names
new.sample.probs.df <- data.frame(tab)
colnames(new.sample.probs.df) <- c("mutations", "frequencies")
#joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
new.sample.df <- left_join_NA(df, new.sample.probs.df, by = "mutations")

#Add pseudocount to column of probabilities 
pole.mmr.ffpe.matrix<- as.matrix(new.sample.df$frequencies) + 0.0001
rownames(pole.mmr.ffpe.matrix)= mutations

plot_96_profile(pole.mmr.ffpe.matrix)
```

##Run classifier function on the FFPE-added sample, remove FFPE mutations
```{r}
classification.df <- get_classification_df(list(pole.sample.vector, mmr.sample.vector, ffpe.sample.vector), c("POLE", "MMR", "FFPE"), list(pole.signature, mmr.signature, ffpe.signature))

ffpe.removed <- classification.df %>%
  filter(classify != "FFPE")

reconstructed.vector <- ffpe.removed$mutations
```

```{r}
#transform sample.vector into a matrix of probabilities
tab <- table(reconstructed.vector)/sum(table(reconstructed.vector))
##converts table of probabilities to a data frame and defines column names
reconstructed.probs.df <- data.frame(tab)
colnames(reconstructed.probs.df) <- c("mutations", "frequencies")
#joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
reconstructed.df <- left_join_NA(df, reconstructed.probs.df, by = "mutations")

#Add pseudocount to column of probabilities 
reconstructed.matrix<- as.matrix(reconstructed.df$frequencies) + 0.0001
rownames(reconstructed.matrix)= mutations
```

```{r}
plot_96_profile(reconstructed.matrix)
```

```{r}
cos_sim_matrix(pole.mmr.matrix, reconstructed.matrix)
```