### 2021-07-12 plots.Rmd
Added point labels to average misclassification rate vs. cosine similarity to FFPE scatterplot, started working on AUC vs cosine similarity to FFPE scatterplot

---
title: "Plots"
author: "Audrey Mitchell"
date: "6/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(NMF)
library(tidyverse)
library(ggrepel)
library(ROCR)
source("bayes_function.R")
source("get_classification_df.R")
source("simulation.R")
```

```{r}
#load FFPE signature as a matrix
ffpe.signature <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))
```

```{r}
#create vector of the 96 mutation types
mutations <- rownames(ffpe.signature) 
```

```{r}
#Load COSMIC signatures (version 3)
cosmic.signatures <- get_known_signatures(muttype = "snv")

#get MMR signature (SBS6)
cosmic.6 <- as.matrix(cosmic.signatures[,"SBS6"])
rownames(cosmic.6) <- mutations
```

#ROC Curve for Sample of 50 MMR mutations and 50 FFPE mutations

```{r}
mmr.sample <-create_signature_sample_vector(cosmic.6, 50)

ffpe.sample <- create_signature_sample_vector(ffpe.signature, 50)

class.df <- get_classification_df(list(mmr.sample, ffpe.sample), c("MMR", "FFPE"), list(cosmic.6, ffpe.signature))
```

```{r}
class.df2 <- class.df %>%
  mutate(mmr.indicator = ifelse(truth == "MMR", 1, 0))
```

```{r}
predictions <- class.df2$MMR
labels <- class.df2$mmr.indicator
pred <- prediction(predictions, labels)
pred
```

```{r}
perf <- performance(pred, "tpr", "fpr")
perf
```

```{r}
plot(perf,
     avg="threshold")
```

```{r}
auc.perf = performance(pred, measure = "auc")
auc.perf@y.values[[1]][1]
```

#Plotting Reconstructed Profiles
- Create a mutational profile without FFPE -> plot mutational profile
- Add FFPE -> plot mutational profile
- Run get_classification_df() function on the FFPE-added mutational profile
- Filter out all the mutations that our function classifies as FFPE (Pr(FFPE) > .5)
- Calculate cosine similarity between original mutational profile and reconstructed

##Create a mutational profile without FFPE
```{r}
#load COSMIC v3 
cosmic.v3 <- get_known_signatures(muttype = "snv")

#create POLE signature matrix
pole.signature <- as.matrix(cosmic.v3[,"SBS10b"])
rownames(pole.signature) <- mutations

#create POLE sample vector
pole.sample.vector <- create_signature_sample_vector(pole.signature, 100)

#create MMR signature matrix
mmr.signature <- as.matrix(cosmic.v3[,"SBS6"])
rownames(mmr.signature) <- mutations

#create MMR sample vector
mmr.sample.vector <- create_signature_sample_vector(mmr.signature, 100)

#Combine POLE and MMR vector
sample.vector <- c(pole.sample.vector, mmr.sample.vector)

#transform sample.vector into a matrix of probabilities
tab <- table(sample.vector)/sum(table(sample.vector))
##converts table of probabilities to a data frame and defines column names
sample.probs.df <- data.frame(tab)
colnames(sample.probs.df) <- c("mutations", "frequencies")

#defines a function that joins two data frames and replaces NA with 0
left_join_NA <- function(x, y, by) {
  left_join(x = x, y = y, by = by) %>% 
  mutate_each(funs(replace(., which(is.na(.)), 0)))
}

#creates a data.frame of mutation types to join with sample probabilities
df <- data.frame(mutations)

#joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
sample.df <- left_join_NA(df, sample.probs.df, by = "mutations")

#Add pseudocount to column of probabilities 
pole.mmr.matrix<- as.matrix(sample.df$frequencies) + 0.0001
rownames(pole.mmr.matrix)= mutations

plot_96_profile(pole.mmr.matrix)
```

#Add FFPE to the sample
```{r}
ffpe.sample.vector <- create_signature_sample_vector(ffpe.signature, 50)

new.sample.vector <- c(sample.vector, ffpe.sample.vector)

#transform sample.vector into a matrix of probabilities
tab <- table(new.sample.vector)/sum(table(new.sample.vector))
##converts table of probabilities to a data frame and defines column names
new.sample.probs.df <- data.frame(tab)
colnames(new.sample.probs.df) <- c("mutations", "frequencies")
#joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
new.sample.df <- left_join_NA(df, new.sample.probs.df, by = "mutations")

#Add pseudocount to column of probabilities 
pole.mmr.ffpe.matrix<- as.matrix(new.sample.df$frequencies) + 0.0001
rownames(pole.mmr.ffpe.matrix)= mutations

plot_96_profile(pole.mmr.ffpe.matrix)
```

##Run classifier function on the FFPE-added sample, remove FFPE mutations
```{r}
classification.df <- get_classification_df(list(pole.sample.vector, mmr.sample.vector, ffpe.sample.vector), c("POLE", "MMR", "FFPE"), list(pole.signature, mmr.signature, ffpe.signature))

ffpe.removed <- classification.df %>%
  filter(classify != "FFPE")

reconstructed.vector <- ffpe.removed$mutations
```

```{r}
#transform sample.vector into a matrix of probabilities
tab <- table(reconstructed.vector)/sum(table(reconstructed.vector))
##converts table of probabilities to a data frame and defines column names
reconstructed.probs.df <- data.frame(tab)
colnames(reconstructed.probs.df) <- c("mutations", "frequencies")
#joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
reconstructed.df <- left_join_NA(df, reconstructed.probs.df, by = "mutations")

#Add pseudocount to column of probabilities 
reconstructed.matrix<- as.matrix(reconstructed.df$frequencies) + 0.0001
rownames(reconstructed.matrix)= mutations
```

```{r}
plot_96_profile(reconstructed.matrix)
```

```{r}
cos_sim_matrix(pole.mmr.matrix, reconstructed.matrix)
```

#Scatterplot of misclassification rate vs. cosine similarity of COSMIC signature to FFPE signature

```{r}
#Create vector of the signature names
signature.names <- colnames(cosmic.signatures)
```

```{r}
#Create list of signature matrices with assigned rownames
signature.matrix.list <- list()
for (i in 1:60){
  sig <- cosmic.signatures[,i]
  sig.mat <- as.matrix(sig)
  rownames(sig.mat) <- mutations
  colnames(sig.mat) <- signature.names[i]
  signature.matrix.list[[i]] <- sig.mat
}
```

```{r}
#outer for loop: loop through each COSMIC v3 signature
average.misclassification.rates <- c()
average.auc.values <- c()
for (x in 1:60){
  signature.misclass.vec <- c()
  signature.auc.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 180 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],180)
    #Create FFPE sample vector of 20 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 20)
    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    misclassification.rate <- mean(classification.df$misclassification)
    signature.misclass.vec[y] <- misclassification.rate
    class.df.colnames <- colnames(classification.df)
    sig.name <- class.df.colnames[3]
    classification.df.2 <- classification.df %>%
      mutate(signature.indicator = ifelse(classify == sig.name, 1, 0))
    pred <- prediction(classification.df.2[3], classification.df.2$signature.indicator)
    auc.perf <- performance(pred, measure = "auc")
    auc.value <- auc.perf@y.values[[1]][1]
    signature.auc.vec[y] <- auc.value
  }
  average.signature.misclass <- mean(signature.misclass.vec)
  average.misclassification.rates[x] <- average.signature.misclass
  average.signature.auc <- mean(signature.auc.vec)
  average.auc.values[x] <- average.signature.auc
}
```

```{r}
#Example of loop for SBS5: all mutations are classified as SBS5 -> problem with ROCR
sample.vec <- create_signature_sample_vector(signature.matrix.list[[5]],180)
    #Create FFPE sample vector of 20 mutations
ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 20)
classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[5],"FFPE"), list(signature.matrix.list[[5]], ffpe.signature))
misclassification.rate <- mean(classification.df$misclassification)
signature.misclass.vec[10] <- misclassification.rate
class.df.colnames <- colnames(classification.df)
sig.name <- class.df.colnames[3]
classification.df.2 <- classification.df %>%
  mutate(signature.indicator = ifelse(classify == sig.name, 1, 0))
pred <- prediction(classification.df.2[3], classification.df.2$signature.indicator)
    auc.perf <- performance(pred, measure = "auc")
    auc.value <- auc.perf@y.values[[1]][1]
    signature.auc.vec[10] <- auc.value
```

```{r}
#create vector of cosine similarities to the FFPE signature
ffpe.cosmic.cos.sims <- c()
for (i in 1:60){
  cos.sim <- cos_sim_matrix(signature.matrix.list[[i]], ffpe.signature)
  ffpe.cosmic.cos.sims[i] <- cos.sim[1,1]
}
```

```{r}
cos.sim.misclass.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.misclassification.rates)
colnames(cos.sim.misclass.data) <- c("signature", "cosine_similarity", "average_misclassification")
```

```{r}
ggplot(cos.sim.misclass.data, aes(x = cosine_similarity, y = average_misclassification))+
  geom_point()+
  geom_smooth(method = "lm")+
  labs(title = "Average misclassification rate vs. cosine similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average misclassification rate (10 iterations)")+
  scale_x_log10()+
  geom_label_repel(aes(label = signature), max.overlaps = 7, label.size =   0.1)
```