### 2021-07-12 firevat/firevat_testing.Rmd

---
title: "firevat_testing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(FIREVAT)
library(bedr)
library(stringr)
library(MutationalPatterns)

source("../simulation.R")
```

```{r}
cosmic.sigs <- get_known_signatures()
cosmic.sig6 <- as.matrix(cosmic.sigs[,6])
ffpe.sig <- as.matrix(load_old_mutational_matrix("../supplied_results/ffpe.signature.txt")) # Doesn't work

df.sigs <- data.frame(GetPCAWGMutSigs()[1:3], cosmic.sigs, ffpe.sig)
rownames(df.sigs) <- seq(1, 96)
```

```{r}
# Output directory
output.dir <- "./" # assign this path

# VCF file
sample.vcf.file <- system.file("extdata", "DCC_PCAWG_Cell_Lines_HCC1954.vcf", package = "FIREVAT")
sample.vcf.file <- "../sample3.vcf"
sample.vcf.file <- "./short-example.vcf"
sample.vcf.file <- "../sample-combined.vcf"
sample.vcf.file <- system.file("extdata", "intestine1-sample.vcf", package = "MutationalPatterns")

# Configuration file
config.file <- system.file("config", "PCAWG_DKFZ_Cell_Line_Filtering_Params.json", package = "FIREVAT")
config.file <- "../custom_config.json"

colnames(df.sigs)[4:length(colnames(df.sigs))]
```

```{r}
run <- function(signature = "SBS6", default = 0) {

  if (default == 0) {
    # Run FIREVAT
    results <- RunFIREVAT(vcf.file = sample.vcf.file,
                          vcf.file.genome = 'hg38', # for mouse variants: 'mm10'
                          config.file = config.file,
                          df.ref.mut.sigs = df.sigs,
                          target.mut.sigs = c(signature, "ffpe.sig"),
                          sequencing.artifact.mut.sigs = c("ffpe.sig"),
                          output.dir = output.dir,
                          objective.fn = Default.Obj.Fn,
                          num.cores = 2,
                          ga.pop.size = 100,
                          ga.max.iter = 5,
                          ga.run = 5,
                          perform.strand.bias.analysis = FALSE,
                          annotate = FALSE)
  } else if (default == 1) {
    # Run FIREVAT
    results <- RunFIREVAT(vcf.file = sample.vcf.file,
                          vcf.file.genome = 'hg19', # for mouse variants: 'mm10'
                          config.file = config.file,
                          df.ref.mut.sigs = GetPCAWGMutSigs(),
                          target.mut.sigs = GetPCAWGMutSigsNames(),
                          sequencing.artifact.mut.sigs = PCAWG.All.Sequencing.Artifact.Signatures,
                          output.dir = output.dir,
                          objective.fn = Default.Obj.Fn,
                          num.cores = 2,
                          ga.pop.size = 100,
                          ga.max.iter = 5,
                          ga.run = 5,
                          perform.strand.bias.analysis = FALSE,
                          annotate = FALSE)
  }
}
```

```{r}
run("SBS4")
```

```{r}
GetPCAWGMutSigs()
GetPCAWGMutSigsNames()
PCAWG.All.Sequencing.Artifact.Signatures
```

```{r}
vcf.obj <- ReadVCF(sample.vcf.file)

mutalisk.results <- RunMutalisk(vcf.obj = vcf.obj,
                                df.ref.mut.sigs = df.sigs,
                                target.mut.sigs = c("SBS6", "ffpe.sig"),
                                n.sample = 2,
                                n.iter = 10)

g <- PlotMutaliskResults(
    mutalisk.results = mutalisk.results,
    signatures = unique(mutalisk.results$identified.mut.sigs),
    df.ref.sigs.groups.colors = GetPCAWGMutSigsEtiologiesColors(),
    trinuc.max.y = 1,
    trinuc.min.y = 0,
    mut.type.max.y = 1,
    title = "")

print(g$f1)

```