### 2021-07-21 plots.Rmd

---
title: "Plots"
author: "Audrey Mitchell"
date: "6/30/2021"
output: pdf_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(NMF)
library(MutationalPatterns)
library(tidyverse)
library(ggrepel)
library(ROCR)
source("bayes_function.R")
source("get_classification_df.R")
source("simulation.R")
```

```{r}
#load FFPE signature as a matrix
ffpe.signature <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))
```

```{r}
#create vector of the 96 mutation types
mutations <- rownames(ffpe.signature) 
```

```{r}
#Load COSMIC signatures (version 3)
cosmic.signatures <- get_known_signatures(muttype = "snv")

#get MMR signature (SBS6)
cosmic.6 <- as.matrix(cosmic.signatures[,"SBS6"])
rownames(cosmic.6) <- mutations
```

#ROC Curve for Sample of 50 MMR mutations and 50 FFPE mutations

```{r}
mmr.sample <-create_signature_sample_vector(cosmic.6, 50)

ffpe.sample <- create_signature_sample_vector(ffpe.signature, 50)

class.df <- get_classification_df(list(mmr.sample, ffpe.sample), c("MMR", "FFPE"), list(cosmic.6, ffpe.signature))
```

```{r}
class.df2 <- class.df %>%
  mutate(mmr.indicator = ifelse(truth == "MMR", 1, 0))
```

```{r}
predictions <- class.df2$MMR
labels <- class.df2$mmr.indicator
pred <- prediction(predictions, labels)
pred
```

```{r}
perf <- performance(pred, "tpr", "fpr")
perf
```

```{r}
plot(perf,
     avg="threshold")
```

```{r}
auc.perf = performance(pred, measure = "auc")
auc.perf@y.values[[1]][1]
```

#Plotting Reconstructed Profiles
- Create a mutational profile without FFPE -> plot mutational profile
- Add FFPE -> plot mutational profile
- Run get_classification_df() function on the FFPE-added mutational profile
- Filter out all the mutations that our function classifies as FFPE (Pr(FFPE) > .5)
- Calculate cosine similarity between original mutational profile and reconstructed

##Create a mutational profile without FFPE
```{r}
#load COSMIC v3 
cosmic.v3 <- get_known_signatures(muttype = "snv")

#create POLE signature matrix
pole.signature <- as.matrix(cosmic.v3[,"SBS10b"])
rownames(pole.signature) <- mutations

#create POLE sample vector
pole.sample.vector <- create_signature_sample_vector(pole.signature, 100)

#create MMR signature matrix
mmr.signature <- as.matrix(cosmic.v3[,"SBS6"])
rownames(mmr.signature) <- mutations

#create MMR sample vector
mmr.sample.vector <- create_signature_sample_vector(mmr.signature, 100)

#Combine POLE and MMR vector
sample.vector <- c(pole.sample.vector, mmr.sample.vector)

#transform sample.vector into a matrix of probabilities
tab <- table(sample.vector)/sum(table(sample.vector))
##converts table of probabilities to a data frame and defines column names
sample.probs.df <- data.frame(tab)
colnames(sample.probs.df) <- c("mutations", "frequencies")

#defines a function that joins two data frames and replaces NA with 0
left_join_NA <- function(x, y, by) {
  left_join(x = x, y = y, by = by) %>% 
  mutate_each(funs(replace(., which(is.na(.)), 0)))
}

#creates a data.frame of mutation types to join with sample probabilities
df <- data.frame(mutations)

#joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
sample.df <- left_join_NA(df, sample.probs.df, by = "mutations")

#Add pseudocount to column of probabilities 
pole.mmr.matrix<- as.matrix(sample.df$frequencies) + 0.0001
rownames(pole.mmr.matrix)= mutations

plot_96_profile(pole.mmr.matrix)
```

#Add FFPE to the sample
```{r}
ffpe.sample.vector <- create_signature_sample_vector(ffpe.signature, 50)

new.sample.vector <- c(sample.vector, ffpe.sample.vector)

#transform sample.vector into a matrix of probabilities
tab <- table(new.sample.vector)/sum(table(new.sample.vector))
##converts table of probabilities to a data frame and defines column names
new.sample.probs.df <- data.frame(tab)
colnames(new.sample.probs.df) <- c("mutations", "frequencies")
#joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
new.sample.df <- left_join_NA(df, new.sample.probs.df, by = "mutations")

#Add pseudocount to column of probabilities 
pole.mmr.ffpe.matrix<- as.matrix(new.sample.df$frequencies) + 0.0001
rownames(pole.mmr.ffpe.matrix)= mutations

plot_96_profile(pole.mmr.ffpe.matrix)
```

##Run classifier function on the FFPE-added sample, remove FFPE mutations
```{r}
classification.df <- get_classification_df(list(pole.sample.vector, mmr.sample.vector, ffpe.sample.vector), c("POLE", "MMR", "FFPE"), list(pole.signature, mmr.signature, ffpe.signature))

ffpe.removed <- classification.df %>%
  filter(classify != "FFPE")

reconstructed.vector <- ffpe.removed$mutations
```

```{r}
#transform sample.vector into a matrix of probabilities
tab <- table(reconstructed.vector)/sum(table(reconstructed.vector))
##converts table of probabilities to a data frame and defines column names
reconstructed.probs.df <- data.frame(tab)
colnames(reconstructed.probs.df) <- c("mutations", "frequencies")
#joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
reconstructed.df <- left_join_NA(df, reconstructed.probs.df, by = "mutations")

#Add pseudocount to column of probabilities 
reconstructed.matrix<- as.matrix(reconstructed.df$frequencies) + 0.0001
rownames(reconstructed.matrix)= mutations
```

```{r}
plot_96_profile(reconstructed.matrix)
```

```{r}
cos_sim_matrix(pole.mmr.matrix, reconstructed.matrix)
```

#Scatterplot of misclassification rate vs. cosine similarity of COSMIC signature to FFPE signature

```{r}
#Create vector of the signature names
signature.names <- colnames(cosmic.signatures)
```

```{r}
#Create list of signature matrices with assigned rownames
signature.matrix.list <- list()
for (i in 1:60){
  sig <- cosmic.signatures[,i]
  sig.mat <- as.matrix(sig)
  rownames(sig.mat) <- mutations
  colnames(sig.mat) <- signature.names[i]
  signature.matrix.list[[i]] <- sig.mat
}
```

```{r}
#outer for loop: loop through each COSMIC v3 signature
average.misclassification.rates <- c()
for (x in 1:60){
  signature.misclass.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 180 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
    #Create FFPE sample vector of 20 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    misclassification.rate <- mean(classification.df$misclassification)
    signature.misclass.vec[y] <- misclassification.rate
  }
  average.signature.misclass <- mean(signature.misclass.vec)
  average.misclassification.rates[x] <- average.signature.misclass
}
```

```{r}
#create vector of cosine similarities to the FFPE signature
ffpe.cosmic.cos.sims <- c()
for (i in 1:60){
  cos.sim <- cos_sim_matrix(signature.matrix.list[[i]], ffpe.signature)
  ffpe.cosmic.cos.sims[i] <- cos.sim[1,1]
}
```

```{r}
cos.sim.misclass.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.misclassification.rates)
colnames(cos.sim.misclass.data) <- c("signature", "cosine_similarity", "average_misclassification")
```

```{r}
model1 <- lm(average_misclassification ~ log(cosine_similarity), data = cos.sim.misclass.data)

summary(model1)

model1_summary <- summary(model1)

cos.sim.misclass.data$residuals <- model1_summary$residuals

cos.sim.misclass.outliers <- cos.sim.misclass.data %>%
  filter(residuals >= quantile(cos.sim.misclass.data$residuals, 0.90) | 
         residuals <= quantile(cos.sim.misclass.data$residuals, 0.1))
```

```{r}
ggplot(cos.sim.misclass.data, aes(x = cosine_similarity, y = average_misclassification))+
  geom_point()+
  geom_smooth(method = "lm")+
  labs(title = "Average misclassification rate vs. cosine similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average misclassification rate (10 iterations)")+
  scale_x_log10()+
  geom_point(data = cos.sim.misclass.outliers, color = "red")+
  geom_label_repel(data = cos.sim.misclass.outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)

```

#Scatterplot of Average AUC vs. Cosine Similarity to FFPE
```{r}
#outer for loop: loop through each COSMIC v3 signature
average.auc.values <- c()
for (x in 1:60){
  signature.auc.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df.1 <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df.1)
    sig.name <- class.df.colnames[3]
    classification.df.2 <- classification.df.1 %>%
      mutate(signature.indicator = ifelse(truth == sig.name, 1, 0))
    predictions <- classification.df.2[3]
    labels <- classification.df.2$signature.indicator
    pred <- prediction(predictions, labels)
    auc.perf <- performance(pred, measure = "auc")
    signature.auc.vec[y] <- auc.perf@y.values[[1]][1]
  }
  average.signature.auc <- mean(signature.auc.vec)
  average.auc.values[x] <- average.signature.auc
}
```

```{r}
cos.sim.auc.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.auc.values)
colnames(cos.sim.auc.data) <- c("signature", "cosine_similarity", "average_auc")
```

```{r}
model2 <- lm(average_auc ~ cosine_similarity, data = cos.sim.auc.data)

summary(model2)

model2_summary <- summary(model2)

cos.sim.auc.data$residuals <- model2_summary$residuals

cos.sim.auc.outliers <- cos.sim.auc.data %>%
  filter(residuals >= quantile(cos.sim.auc.data$residuals, 0.95) | 
         residuals <= quantile(cos.sim.auc.data$residuals, 0.05))
```

```{r}
ggplot(cos.sim.auc.data, aes(x = cosine_similarity, y = average_auc))+
  geom_point()+
  geom_smooth(method = "lm")+
  labs(title = "Average area under curve vs. cosine similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average area under curve (10 iterations)")+
  geom_point(data = cos.sim.auc.outliers, color = "red")+
  geom_label_repel(data = cos.sim.auc.outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)

```

#Varying total number of mutations (Signature 4)

```{r}
signature.4.mat <- as.matrix(cosmic.signatures[,"SBS4"])
rownames(signature.4.mat) <- mutations
plot_96_profile(signature.4.mat)
```

```{r}
average.auc.values <- c()
i = 1
for(total_mutations in seq(100, 1000, by = 50)){
  auc.values <- c()
  for(x in 1:10){
    mutation.number <- total_mutations/2
    sig.4.sample.vec <- create_signature_sample_vector(signature.4.mat, mutation.number)
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, mutation.number)
    classification <- get_classification_df(list(sig.4.sample.vec, ffpe.sample.vec), c("SBS4","FFPE"), list(signature.4.mat, ffpe.signature))
    classification <- classification %>%
      mutate(indicator = ifelse(truth == "SBS4", 1, 0))
    predictions <- classification$SBS4
    labels <- classification$indicator
    pred = prediction(predictions, labels)
    auc.perf <- performance(pred, measure = "auc")
    auc.val <- auc.perf@y.values[[1]][1]
    auc.values[x] <- auc.val
  }
  average.auc.val <- mean(auc.values)
  average.auc.values[i] <- average.auc.val
  i = i + 1

}
```

```{r}
total_mutations <- seq(100, 1000, by = 50)
total_mutations_df <- data.frame(total_mutations, average.auc.values)
colnames(total_mutations_df) <- c("total_mutations", "average_auc")

#plot(total_mutations_df$total_mutations, total_mutations_df$average.auc.values)
```

```{r}
ggplot(data = total_mutations_df, aes(x = total_mutations, y = average_auc))+
  geom_point()+
  geom_smooth(method = "lm", se = TRUE)
```

#Varying total number of mutations (Signature 6)

```{r}
signature.6.mat <- as.matrix(cosmic.signatures[,"SBS6"])
rownames(signature.6.mat) <- mutations
plot_96_profile(signature.6.mat)
```

```{r}
average.auc.values6 <- c()
i = 1
for(total_mutations in seq(100, 1000, by = 50)){
  auc.values6 <- c()
  for(x in 1:10){
    mutation.number <- total_mutations/2
    sig.6.sample.vec <- create_signature_sample_vector(signature.6.mat, mutation.number)
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, mutation.number)
    classification <- get_classification_df(list(sig.6.sample.vec, ffpe.sample.vec), c("SBS6","FFPE"), list(signature.6.mat, ffpe.signature))
    classification <- classification %>%
      mutate(indicator = ifelse(truth == "SBS6", 1, 0))
    predictions <- classification$SBS6
    labels <- classification$indicator
    pred = prediction(predictions, labels)
    auc.perf <- performance(pred, measure = "auc")
    auc.val <- auc.perf@y.values[[1]][1]
    auc.values[x] <- auc.val
  }
  average.auc.val <- mean(auc.values)
  average.auc.values6[i] <- average.auc.val
  i = i + 1

}
```

```{r}
total_mutations6 <- seq(100, 1000, by = 50)
total_mutations_df6 <- data.frame(total_mutations6, average.auc.values6)
colnames(total_mutations_df6) <- c("total_mutations", "average_auc")

#plot(total_mutations_df$total_mutations, total_mutations_df$average.auc.values)
```

```{r}
ggplot(data = total_mutations_df6, aes(x = total_mutations, y = average_auc))+
  geom_point()
  #geom_smooth(method = "lm")
```

#Diagram of Method
- We start with a mutational profile: create a sample of 200 MMR, 200 Tobacco, 100 FFPE, plot the mutational profile

```{r}
mmr.sample <- create_signature_sample_vector(signature.6.mat, 200)
tobacco.sample <- create_signature_sample_vector(signature.4.mat, 200)
ffpe.sample <- create_signature_sample_vector(ffpe.signature, 100)

ex.profile.mat <- vector_to_matrix(c(mmr.sample, tobacco.sample))
rownames(ex.profile.mat) <- mutations

plot_96_profile(ex.profile.mat)+
  labs(title = "Mutational Profile of Original Simulated Sample")+
  theme(aspect.ratio = 1)
```

```{r}
ex.ffpe.profile.mat <- vector_to_matrix(c(mmr.sample, tobacco.sample, ffpe.sample))
rownames(ex.ffpe.profile.mat) <- mutations

plot_96_profile(ex.ffpe.profile.mat)+
  labs(title = "Mutational Profile After Addition of FFPE Mutations")+
  theme(aspect.ratio = 1)
```

- We use extract_signatures or fit_to_signatures from the MutationalPatterns package to determine mutational signatures and their contributions: plot signatures and contributions

```{r}
sig.mat <- cbind(signature.4.mat, signature.6.mat, ffpe.signature)
colnames(sig.mat) <- c("Signature 4", "Signature 6", "FFPE")
fts.res <- fit_to_signatures(ex.ffpe.profile.mat, sig.mat)

plot_contribution(fts.res$contribution, sig.mat, coord_flip = TRUE)+
  labs(title = "Relative Contributions of Signatures")+
  theme(aspect.ratio = 0.3)
```

```{r}
plot_96_profile(sig.mat)+
  labs(title = "Mutational Signatures")
```

- We use Bayes formula to calculate the posterior probability of each signature given each of the 96 mutation types; prior, likelihood, and marginalization are obtained from the mutational signatures and contributions
- We classify which signature each mutation is due to based on the signature with the hightest posterior probability: show part of the classification data frame

```{r}
classifications <- get_classification_df(list(tobacco.sample, mmr.sample, ffpe.sample), c("Signature 4", "Signature 6", "FFPE"), list(signature.4.mat, signature.6.mat, ffpe.signature))

head(classifications)
```

- Mutations classified as FFPE are filtered out, and we obtain a reconstructed profile: plot reconstructed profile

```{r}
filter.ffpe <- classifications %>%
  filter(classify != "FFPE")

recon.vec <- filter.ffpe$mutations
recon.mat <- vector_to_matrix(recon.vec)
rownames(recon.mat) <- mutations

plot_96_profile(recon.mat)+
  labs(title = "Reconstructed Profile After Removing FFPE Classifed Mutations")
```

```{r}
cos_sim_matrix(ex.profile.mat, recon.mat)
```

#Description of Simulation
- Plot original signature
- Plot profile of sample created to match the signature distribution
- cosine similarity of sample to original signature

```{r}
x1 <- plot_96_profile(signature.4.mat)+
  labs(title = "Signature 4")
x2 <- plot_96_profile(signature.6.mat)+
  labs(title = "Signature 6")
tobacco.1000 <- create_signature_sample_matrix(signature.4.mat, 1000)
mmr.1000 <- create_signature_sample_matrix(signature.6.mat, 1000)
x3 <- plot_96_profile(tobacco.1000)+
  labs(title = "Signature 4 Simulated Sample Profile")
x4 <- plot_96_profile(mmr.1000)+
  labs(title = "Signature 6 Simulated Sample Profile")
grid.arrange(x1, x2, x3, x4, nrow = 2)
```

#Profile Reconstruction for MMR
- Create sample of 500 MMR mutations
- Plot mutational profile of sample
- Add 500 FFPE mutations to the sample
- Plot FFPE-added profile
- Run classifier function
- Filter out FFPE mutations
- Plot reconstructed profile
- Calculate the cosine similarity between the original and recosntructed

```{r}
mmr.500 <- create_signature_sample_vector(signature.6.mat, 500)
mmr.mat <- vector_to_matrix(mmr.500)
rownames(mmr.mat) <- mutations

m1 <- plot_96_profile(mmr.mat)+
  labs(title = "Mutational Profile of Signature 6 Simulated Sample")+
  theme(axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        plot.title = element_text(size = 10))
```

```{r}
ffpe.500 <- create_signature_sample_vector(ffpe.signature, 500)
mmr.ffpe.samp <- c(mmr.500, ffpe.500)
mmr.ffpe.mat <- vector_to_matrix(mmr.ffpe.samp)
rownames(mmr.ffpe.mat) <- mutations
m2 <- plot_96_profile(mmr.ffpe.mat)+
  labs(title = "Mutational Profile After Addition of FFPE")+
  theme(axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        plot.title = element_text(size = 10))
```

```{r}
mmr.ffpe.class <- get_classification_df(list(mmr.500, ffpe.500), c("Signature 6", "FFPE"), list(signature.6.mat, ffpe.signature))

mmr.only <- mmr.ffpe.class %>%
  filter(classify != "FFPE")

recon.mmr.vec <- mmr.only$mutations
recon.mmr.matrix <- vector_to_matrix(recon.mmr.vec)
rownames(recon.mmr.matrix) <- mutations

m3 <- plot_96_profile(recon.mmr.matrix)+
  labs(title = "Reconstructed Mutational Profile (FFPE Mutations Removed)")+
  theme(axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        plot.title = element_text(size = 10))
```

```{r}
grid.arrange(m1,m2,m3, nrow = 3)
```

#Profile Reconstruction for Signature 4
- Create sample of 500 Signature 4 mutations
- Plot mutational profile of sample
- Add 500 FFPE mutations to the sample
- Plot FFPE-added profile
- Run classifier function
- Filter out FFPE mutations
- Plot reconstructed profile
- Calculate the cosine similarity between the original and recosntructed

```{r}
tobacco.500 <- create_signature_sample_vector(signature.4.mat, 500)
tobacco.mat <- vector_to_matrix(tobacco.500)

t1 <- plot_96_profile(tobacco.mat)+
  labs(title = "Mutational Profile of Signature 4 Simulated Sample", x = "Dinucleotide Context", y = "Probability")+
  theme(axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        plot.title = element_text(size = 10))
t1
```

```{r}
tobacco.ffpe.vec <- c(tobacco.500, ffpe.500)
tobacco.ffpe.mat <- vector_to_matrix(tobacco.ffpe.vec)
rownames(tobacco.ffpe.mat) <- mutations

t2 <- plot_96_profile(tobacco.ffpe.mat)+
  labs(title = "Mutational Profile After Addition of FFPE", x = "Dinucleotide Context", y = "Probability")+
  theme(axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        plot.title = element_text(size = 10))
```

```{r}
tobacco.ffpe.class <- get_classification_df(list(tobacco.500, ffpe.500), c("Signature 4", "FFPE"), list(signature.4.mat, ffpe.signature))

tobacco.only <- tobacco.ffpe.class %>%
  filter(classify != "FFPE")

recon.tobacco.vec <- tobacco.only$mutations
recon.tobacco.mat <- vector_to_matrix(recon.tobacco.vec)
rownames(recon.tobacco.mat) <- mutations

t3 <- plot_96_profile(recon.tobacco.mat)+
  labs(title = "Reconstructed Mutational Profile (FFPE Mutations Removed)", x = "Dinucleotide Context", y = "Probability")+
  theme(axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        plot.title = element_text(size = 10))
```

```{r}
grid.arrange(t1, t2, t3)
```

#Scatterplot of Cosine similarity between recosntructed signature and original vs. cosine similarity between signature and the FFPE signature
- For each signature, iterate through sample creation and profile reconstruction 10 times, calculate cosine similarity between original and reconstruction
- Take average of 10 cosine similarities, plot against the cosine similarity of the COSMIC signatures to FFPE

```{r}
#outer for loop: loop through each COSMIC v3 signature
average.auc.values <- c()
for (x in 1:60){
  signature.auc.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df.1 <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df.1)
    sig.name <- class.df.colnames[3]
    classification.df.2 <- classification.df.1 %>%
      mutate(signature.indicator = ifelse(truth == sig.name, 1, 0))
    predictions <- classification.df.2[3]
    labels <- classification.df.2$signature.indicator
    pred <- prediction(predictions, labels)
    auc.perf <- performance(pred, measure = "auc")
    signature.auc.vec[y] <- auc.perf@y.values[[1]][1]
  }
  average.signature.auc <- mean(signature.auc.vec)
  average.auc.values[x] <- average.signature.auc
}
```

#Scatterplot of Sensitivity and Specificity vs. Cosine similarity of COSMIC signature to FFPE

```{r}
#outer for loop: loop through each COSMIC v3 signature
average.sensitivity.values <- c()
average.specificity.values <- c()
for (x in 1:60){
  signature.sensitivity.vec <- c()
  signature.specificity.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df)
    sig.name <- class.df.colnames[3]
    #Count number of true positives (for sensititivity)
    sensitivity.df <- classification.df %>%
      filter(truth == sig.name & classify == sig.name)
    #Calculate sensitivity
    sensitivity.val = nrow(sensitivity.df)/100
    #Add to the vector of 10 sensitivity values
    signature.sensitivity.vec[y] = sensitivity.val
    #Count number of true negatives (for specificity)
    specificity.df <- classification.df %>%
      filter(truth == "FFPE" & classify == "FFPE")
    #Calculate specificity
    specificity.val = nrow(specificity.df)/100
    #Add to the vector of 10 specificity values
    signature.specificity.vec[y] = specificity.val
  }
  average.signature.sensitivity <- mean(signature.sensitivity.vec)
  average.sensitivity.values[x] <- average.signature.sensitivity
  average.signature.specificity <- mean(signature.specificity.vec)
  average.specificity.values[x] <- average.signature.specificity
}
```

```{r}
sens.spec.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.sensitivity.values, average.specificity.values)

colnames(sens.spec.data) <- c("signature", "cosine_similarity", "sensitivity", "specificity")

sens.spec.data.2 <- sens.spec.data %>%
  pivot_longer(c(sensitivity, specificity), names_to = "measure", values_to = "value")
```

```{r}
ggplot(data = sens.spec.data.2, aes(x = cosine_similarity, y = value, color = measure))+
  geom_point()+
  geom_smooth(method = lm)+
  labs(title = "Sensitivity & Specificity vs. Cosine Similarity to FFPE Signature", x = "Cosine Similarity of COSMIC Signature to FFPE", y = "Sensitivity and Specificity")
```