### 2022-08-02 Abstract.Rmd
Reformatting of figure 2, replacing AUC for accuracy

---
title: "Abstract Plots"
author: "Audrey Mitchell"
date: "7/21/2021"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(NMF)
library(MutationalPatterns)
library(tidyverse)
library(ggrepel)
library(ROCR)
library(gridExtra)
source("bayes_function.R")
source("get_classification_df.R")
source("simulation.R")
sessionInfo()
```

```{r, include = FALSE}
vector_to_matrix <- function(mutations.vector){
  tab <- table(mutations.vector)/sum(table(mutations.vector))
  #transform table to a data frame and assign column names to mutations and frequencies
  mutations.df <- data.frame(tab)
  colnames(mutations.df) <- c("mutations", "frequencies")

  #defines a function that joins two data frames and replaces NA with 0
  left_join_NA <- function(x, y, by) {
    left_join(x = x, y = y, by = by) %>% 
      mutate_each(funs(replace(., which(is.na(.)), 0)))
  }

  #converts vector of 96 mutation types to a data frame
  mutation.types.df <- data.frame(mutations)
  #joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
  complete.mutations.df <- left_join_NA(mutation.types.df, mutations.df, by = "mutations")

  #converts data frame of mutation types and frequencies to a data frame and assigns rownames
  mutations.matrix <- as.matrix(complete.mutations.df$frequencies)
  rownames(mutations.matrix) = mutations

  return(mutations.matrix)
}
```

```{r, include=FALSE}
#load FFPE signature as a matrix
ffpe.signature <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))

#create vector of the 96 mutation types
mutations <- rownames(ffpe.signature)

#Load COSMIC signatures (version 3)
cosmic.signatures <- get_known_signatures(muttype = "snv")

#Create vector of the signature names
signature.names <- colnames(cosmic.signatures)

#Create list of signature matrices with assigned rownames
signature.matrix.list <- list()
for (i in 1:length(signature.names)){
  sig <- cosmic.signatures[,i]
  sig.mat <- as.matrix(sig)
  rownames(sig.mat) <- mutations
  colnames(sig.mat) <- signature.names[i]
  signature.matrix.list[[i]] <- sig.mat
}

#create vector of cosine similarities to the FFPE signature
ffpe.cosmic.cos.sims <- c()
for (i in 1:length(signature.names)){
  cos.sim <- cos_sim_matrix(signature.matrix.list[[i]], ffpe.signature)
  ffpe.cosmic.cos.sims[i] <- cos.sim[1,1]
}

signature.4.mat <- as.matrix(cosmic.signatures[,"SBS4"])
rownames(signature.4.mat) <- mutations

signature.6.mat <- as.matrix(cosmic.signatures[,"SBS6"])
rownames(signature.6.mat) <- mutations
```

# Diagram of Methods

```{r, include=FALSE}
mmr.sample <- create_signature_sample_vector(signature.6.mat, 200)
tobacco.sample <- create_signature_sample_vector(signature.4.mat, 200)
ffpe.sample <- create_signature_sample_vector(ffpe.signature, 100)

ex.ffpe.profile.mat <- vector_to_matrix(c(mmr.sample, tobacco.sample, ffpe.sample))
rownames(ex.ffpe.profile.mat) <- mutations
colnames(ex.ffpe.profile.mat) <- c("Sample 1")

mut.profile <- plot_96_profile(ex.ffpe.profile.mat)+
  labs(title = "Mutational Profile", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

mut.profile

#ggsave("mutprofile.png", plot = mut.profile , width = 175, height = 108, units = "mm")
```

```{r, include=FALSE}
sig.mat <- cbind(signature.4.mat, signature.6.mat, ffpe.signature)
colnames(sig.mat) <- c("Signature 4", "Signature 6", "FFPE")
fts.res <- fit_to_signatures(ex.ffpe.profile.mat, sig.mat)

contribution.plot <- plot_contribution(fts.res$contribution, sig.mat, coord_flip = TRUE)+
  labs(title = "Relative Contributions of Signatures")+
  theme(plot.title = element_text(hjust = 0.5))

contribution.plot

#ggsave("contributions.png", plot = contribution.plot , width = 175, height = 54, units = "mm")
```

```{r, include=FALSE}
signatures.plot <- plot_96_profile(sig.mat)+
  labs(title = "Mutational Signatures", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

signatures.plot

#ggsave("signatures.png", plot = signatures.plot , width = 175, height = 108, units = "mm")
```

# Diagram of Simulating Data

```{r, include=FALSE}
signature4.plot <- plot_96_profile(signature.4.mat)+
  labs(title = "Signature 4", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

#ggsave("signature4.png", plot = signature4.plot , width = 175, height = 81, units = "mm")
```

```{r, include=FALSE}
ffpe.plot <- plot_96_profile(ffpe.signature)+
  labs(title = "FFPE Signature", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

#ggsave("ffpe.png", plot = ffpe.plot , width = 175, height = 81, units = "mm")
```

```{r, include=FALSE}
sample.4 <- create_signature_sample_vector(signature.4.mat, 500)
sample.ffpe <- create_signature_sample_vector(ffpe.signature, 500)

ex.sim.sample.vec <- c(sample.4, sample.ffpe)
ex.sim.sample <- vector_to_matrix(ex.sim.sample.vec)

sampleprof.plot <- plot_96_profile(ex.sim.sample)+
  labs(title = "Simulated Sample Profile", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

sampleprof.plot

#ggsave("sampleprofile.png", plot = sampleprof.plot , width = 175, height = 108, units = "mm")
```

# Reconstructed Profiles of Signatures 4 and 6

```{r, include=FALSE}
set.seed(500)
#Create Signature 4 Sample
sig4.500 <- create_signature_sample_vector(signature.4.mat, 500)
sig4.profile <- vector_to_matrix(sig4.500)

sig4.profile.plot <- plot_96_profile(sig4.profile)+
  labs(title = "Signature 4 Simulated Sample", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

sig4.profile.plot
```

```{r, include=FALSE}
#Add FFPE Noise
addffpe.500 <- create_signature_sample_vector(ffpe.signature, 500)

sig4.ffpe <- vector_to_matrix(c(sig4.500, addffpe.500))

sig4.ffpe.plot <- plot_96_profile(sig4.ffpe)+
  labs(title = "Simulated Sample with FFPE Noise", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r, include=FALSE}
#Classify and Filter
sig4.ffpe.class <- get_classification_df(list(sig4.500, addffpe.500), c("SBS4", "FFPE"), list(signature.4.mat, ffpe.signature))

sig4.ffpe.filtered <- sig4.ffpe.class %>%
  filter(classify != "FFPE")

sig4.recon.matrix <- vector_to_matrix(sig4.ffpe.filtered$mutations)

sig4.recon.plot <- plot_96_profile(sig4.recon.matrix)+
  labs(title = "Signature 4 Reconstructed", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

sig4.recon.plot
```

```{r, include=FALSE}
#Calculate Sensitivity, Specificity, and Misclassification rate

sig4.sensitivity <- sig4.ffpe.class %>%
  filter(truth == "SBS4" & classify == "SBS4")

nrow(sig4.sensitivity)/500

sig4.specificity <- sig4.ffpe.class %>%
  filter(truth == "FFPE" & classify == "FFPE")

nrow(sig4.specificity)/500

mean(sig4.ffpe.class$misclassification)
```

```{r, include=FALSE}
#Arrange 3 plots
sig4.reconstruction <- grid.arrange(sig4.profile.plot, sig4.ffpe.plot, sig4.recon.plot)

#ggsave("sig4reconstruction.png", plot = sig4.reconstruction , width = 175, height = 243, units = "mm")
```

```{r, include=FALSE}
#Create Signature 6 Sample
sig6.500 <- create_signature_sample_vector(signature.6.mat, 500)
sig6.profile <- vector_to_matrix(sig6.500)

sig6.profile.plot <- plot_96_profile(sig6.profile)+
  labs(title = "Signature 6 Simulated Sample", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

sig6.profile.plot
```

```{r, include=FALSE}
#Add FFPE Noise
sig6.ffpe <- vector_to_matrix(c(sig6.500, addffpe.500))

sig6.ffpe.plot <- plot_96_profile(sig6.ffpe)+
  labs(title = "Simulated Sample with FFPE Noise", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

sig6.ffpe.plot
```

```{r, include=FALSE}
#Classify and Filter
sig6.ffpe.class <- get_classification_df(list(sig6.500, addffpe.500), c("SBS6", "FFPE"), list(signature.6.mat, ffpe.signature))

sig6.ffpe.filtered <- sig6.ffpe.class %>%
  filter(classify != "FFPE")

sig6.recon.matrix <- vector_to_matrix(sig6.ffpe.filtered$mutations)

sig6.recon.plot <- plot_96_profile(sig6.recon.matrix)+
  labs(title = "Signature 6 Reconstructed", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

sig6.recon.plot
```

```{r, include=FALSE}
#Calculate Sensitivity, specificity, and misclassification rate

sig6.sensitivity <- sig6.ffpe.class %>%
  filter(truth == "SBS6" & classify == "SBS6")

nrow(sig6.sensitivity)/500

sig6.specificity <- sig6.ffpe.class %>%
  filter(truth == "FFPE" & classify == "FFPE")

nrow(sig6.specificity)/500

mean(sig6.ffpe.class$misclassification)
```

```{r, include=FALSE}
#Arrange 3 plots
sig6.reconstruction <- grid.arrange(sig6.profile.plot, sig6.ffpe.plot, sig6.recon.plot)

#ggsave("sig6reconstruction.png", plot = sig6.reconstruction , width = 175, height = 243, units = "mm")
```

# ROC curve

```{r, include=FALSE}
sample.sig6 <- create_signature_sample_vector(signature.6.mat, 500)
sig6.truth <- rep("MMR", 500)
sample.ffpe <- create_signature_sample_vector(ffpe.signature, 500)
ffpe.truth <- rep("FFPE", 500)
MMR.ROC <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"),
                              list(signature.matrix.list[[6]], ffpe.signature))

sample.sig4 <- create_signature_sample_vector(signature.4.mat, 500)
sig4.truth <- rep("Tobacco", 500)
tob.ROC <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"),
                              list(signature.matrix.list[[4]], ffpe.signature))
```

```{r, include=FALSE}
MMR.ROC <- MMR.ROC %>%
  #create binary labels
  mutate(truthMMR = ifelse(truth == "MMR",1,0))

tob.ROC <- tob.ROC %>%
  mutate(truthTob = ifelse(truth == "Tobacco",1,0))

#provide predictions and labels
MMR.pred <- prediction(MMR.ROC$MMR, MMR.ROC$truthMMR)
mmr.auc <-  performance(MMR.pred, measure = "auc")
mmr.auc@y.values[[1]][1] #AUC for SBS6
MMR.perf <- performance(MMR.pred, measure = "tpr", x.measure = "fpr")

tob.pred <- prediction(tob.ROC$Tobacco, tob.ROC$truthTob)
tob.auc <-  performance(tob.pred, measure = "auc")
tob.auc@y.values[[1]][1] #AUC for SBS4

tob.perf <- performance(tob.pred, measure = "tpr", x.measure = "fpr")

png("MMRtobROC.png", res = 300, width = 131.25, height = 108, units = "mm")

plot(MMR.perf, print.cutoffs.at=0.5, text.adj=c(-0.2,1.7),lty = 3) #ROC
plot(tob.perf, print.cutoffs.at=0.5, text.adj=c(-0.2,1.7), 
     add = TRUE, lty = 2)
abline(a=0, b= 1)
legend("bottomright", c("SBS4", "SBS6"),
       lty = 2:3)

dev.off()
```

# Accuracy vs Cosine similarity to FFPE

We first create a function that mixes mutations from FFPE and another given mutational signature and calls the bayes classifier `get_classification_df`

```{r}
test_mix_mutations <- function (cosmic_idx, num_cosmic, num_ffpe) {

  # Create baseline mutations
  baseline_sig <- signature.matrix.list[[cosmic_idx]]
  baseline_name <- signature.names[cosmic_idx]
  baseline_mut_vec <- create_signature_sample_vector(baseline_sig, num_cosmic)

  # Create FFPE mutations
  ffpe_mut_vec <- create_signature_sample_vector(ffpe.signature, num_ffpe)

  # Call excerno classification and format as a tibble
  get_classification_df (list(baseline_mut_vec, ffpe_mut_vec), 
                          c(baseline_name,"FFPE"), 
                          list(baseline_sig, ffpe.signature)) %>% 
    as_tibble()
}
```

And we try `test_mix_mutations` and plot the results to check it is working:

```{r}
set.seed(12345)
test_tbl <- test_mix_mutations(1,500,500)

test_tbl %>%
  select (mutations, truth, SBS1, FFPE) %>%
  pivot_longer(c(SBS1, FFPE)) %>%
  ggplot(aes(x=truth, y=value, fill=name)) +
    geom_boxplot()

```

Now let's create a function that calculates the performance metrics (sensitivity, specificity, accuracy) on the output of `test_mix_mutations`

```{r}
calc_metrics_mix <- function (cosmic_idx, num_cosmic, num_ffpe) {

  baseline_name <- signature.names[cosmic_idx]
  test_tbl <- test_mix_mutations(cosmic_idx,num_cosmic,num_ffpe)

  tp <- test_tbl %>%
    filter(truth == baseline_name & classify == baseline_name) %>%
    count() %>%
    pull(n)

  fp <- test_tbl %>%
    filter(truth != baseline_name & classify == baseline_name) %>%
    count() %>%
    pull(n)

  fn <- test_tbl %>%
    filter(truth == baseline_name & classify == "FFPE") %>%
    count() %>%
    pull(n)

  tn <- test_tbl %>%
    filter(truth == "FFPE" & classify == "FFPE") %>%
    count() %>%
    pull(n)

  tibble(
    Sensitivity = tp / (tp +fn),
    Specificity = tn / (tp + fn),
    Accuracy = (tp +tn) /(tp+tn+fp+fn))
}

```

And we test our method:

```{r}
set.seed(12345)

calc_metrics_mix(1,500,500)
calc_metrics_mix(4,500,500)
calc_metrics_mix(6,500,500)
```

We create a function that calculates the average metrics over a fixed number of simulations `num_iter` and test it

```{r}
calc_metrics_mix_num <- function (cosmic_idx, num_iter, num_cosmic, num_ffpe) {

  baseline_name <- signature.names[cosmic_idx]

  1:num_iter %>% 
    map_dfr(~calc_metrics_mix(cosmic_idx,num_cosmic,num_ffpe))  %>%
    pivot_longer(1:3) %>% 
    group_by(name) %>%
    summarize(mean = mean(value)) %>%
    mutate(sbs = baseline_name)

}

calc_metrics_mix_num(1,10,500,500)
calc_metrics_mix_num(4,10,500,500)
calc_metrics_mix_num(6,10,500,500)
```

Finally we are ready to generate the data over all of COSMIC mutational signatures, using 10 iterations and 500 mutations for FFPE and 500 mutations from the baseline COSMIC signature

```{r}
sbs_num = length(signature.names)
base_num = 500
ffpe_num = 500
reps = 10

set.seed(123456)
sim_tbl <- 1:sbs_num %>%
  map_dfr(~calc_metrics_mix_num(.x, reps, base_num, ffpe_num))
```

First we summarize the metrics across all the COSMIC mutational signatures

```{r}
library(kableExtra)

sim_tbl <- sim_tbl %>%
  mutate (name = factor (name, 
                         levels = c("Sensitivity", "Specificity","Accuracy"))) 
sim_tbl %>%
  ggplot(aes(mean))+
  geom_histogram(binwidth = 0.025)+
  facet_grid(.~name) +
  labs(x="Metric across all COSMIC signatures")

ggsave("paper_figures/supp3.png", dpi = 300, width = 175, height = 75, units = "mm")

summary_tbl <- sim_tbl %>%
  group_by(name) %>%
  summarize (median = median (mean),
           IQR = IQR(mean)) %>%
  mutate_if(is.numeric, format, digits=2)

summary_tbl %>%
  kbl() %>%
  kable_styling()
```

Now we explore the relationship of the performance metrics against the similarity between the baseline signature and FFPE using a figure

```{r}
library(broom)
sim_tbl <- sim_tbl %>%
  mutate (ffpe_cos = rep(ffpe.cosmic.cos.sims, each=3))

sim_tbl %>%
  ggplot(aes(x=ffpe_cos, mean)) +
  geom_point()+
  geom_smooth(method="lm", formula = "y~x")+
  facet_grid(~name)

```

And using a linear model

```{r}
lm_metric<- lm(mean ~ name:ffpe_cos, data = sim_tbl)

tidy(lm_metric) %>%
  kbl() %>%
  kable_styling()

glance(lm_metric) %>%
  kbl() %>%
  kable_styling()
```

We define a function to represent the mutational signatures that fall below a threshold for a particular metric.

```{r}
plot_metrics_excerno <- function (tbl, metric, threshold) {

  # Subset the table to the specific metric  
  new_tbl <- filter(tbl, name==metric) 
  # Identify signature that fall below the threshold
  outliers <- filter( new_tbl, mean < threshold)
  # Extract signatures SBS4 and SBS6 as a fixe reference in the graph
  fixed_sbs <- filter( new_tbl, sbs == "SBS4" | sbs == "SBS6")

  ggplot(new_tbl, aes(x = ffpe_cos, y = mean))+
    geom_point()+
    geom_smooth(method = "lm", se=FALSE)+
    geom_point(data = outliers, color = "red")+
    geom_point(data = fixed_sbs, color = "red")+
    geom_hline(yintercept=threshold, linetype="dashed", color = "red")+
    geom_label_repel(data=outliers, 
                   aes(label = sbs),
                   max.overlaps = 10, 
                   label.size = .1)+
    geom_label_repel(data=fixed_sbs, 
                   aes(label = sbs),
                   max.overlaps = 10, 
                   label.size =   0.1) +
   labs(x = "Cosine similarity against FFPE signature", 
         y = metric )
}
```

And we plot all the metrics using a threshold of 80%

```{r}
plot_metrics_excerno(sim_tbl, "Accuracy", 0.8)
ggsave("paper_figures/fig2b.png", dpi = 300, width = 175, height = 108, units = "mm")

plot_metrics_excerno(sim_tbl, "Sensitivity", 0.8)
ggsave("paper_figures/supp6.png", dpi = 300, width = 175, height = 108, units = "mm")

plot_metrics_excerno(sim_tbl, "Specificity", 0.8)
ggsave("paper_figures/supp7.png", dpi = 300, width = 175, height = 108, units = "mm")
```

And we obtain the linear model for each metric separately and print their information

```{r}
print_lm_coefs <- function (tbl, metric) {
  new_tbl <-  filter (tbl, name==metric)
  lm_metrics<- lm(mean ~ ffpe_cos, data = new_tbl)

  tidy(lm_metrics) %>%
    mutate_if(is.numeric, format, digits=2)%>%
    kbl() %>%
    kable_styling()
}

print_lm_fit <- function (tbl, metric) {
  new_tbl <-  filter (tbl, name==metric)
  lm_metrics<- lm(mean ~ ffpe_cos, data = new_tbl)

  glance(lm_metrics) %>%
    mutate_if(is.numeric, format, digits=2)%>%
    kbl() %>%
    kable_styling()
}

print_lm_coefs(sim_tbl, "Accuracy")
print_lm_fit(sim_tbl, "Accuracy")

print_lm_coefs(sim_tbl, "Sensitivity")
print_lm_fit(sim_tbl, "Sensitivity")

print_lm_coefs(sim_tbl, "Specificity")
print_lm_fit(sim_tbl, "Specificity")

```

# Scatterplot of Average AUC vs. Cosine Similarity to FFPE

```{r, include=FALSE}
set.seed(500)
#outer for loop: loop through each COSMIC v3 signature
average.auc.values <- c()
for (x in 1:length(signature.names)){
  signature.auc.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],500)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 500)
    classification.df.1 <- 
      get_classification_df(list(sample.vec, ffpe.sample.vec), 
                            c(signature.names[x],"FFPE"), 
                            list(signature.matrix.list[[x]], ffpe.signature))

    class.df.colnames <- colnames(classification.df.1)
    sig.name <- class.df.colnames[3]
    classification.df.2 <- classification.df.1 %>%
      mutate(signature.indicator = ifelse(truth == sig.name, 1, 0))
    predictions <- classification.df.2[3]
    labels <- classification.df.2$signature.indicator
    pred <- prediction(predictions, labels)
    auc.perf <- performance(pred, measure = "auc")
    signature.auc.vec[y] <- auc.perf@y.values[[1]][1]
  }
  average.signature.auc <- mean(signature.auc.vec)
  average.auc.values[x] <- average.signature.auc
}
```

```{r, include=FALSE}
cos.sim.auc.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.auc.values)
colnames(cos.sim.auc.data) <- c("signature", "cosine_similarity", "average_auc")
```

```{r, include=FALSE}

ggplot(cos.sim.auc.data, aes(x=average_auc)) +
  geom_histogram(color="black", fill="lightblue")+
  labs( x = "Average AUC", y = "Count")
#ggsave("supp.fig.5.png", dpi = 300, width = 175, height = 108, units = "mm")

median(cos.sim.auc.data$average_auc)
IQR(cos.sim.auc.data$average_auc)

model2 <- lm(average_auc ~ cosine_similarity, data = cos.sim.auc.data)

model2_summary <- summary(model2)

cos.sim.auc.data$residuals <- model2_summary$residuals

cos.sim.auc.outliers <- cos.sim.auc.data %>%
  filter(average_auc<0.8)

cos.sim.auc.outliers

cos.sim.auc.outliers2 <- cos.sim.auc.data %>%
  filter(average_auc>=0.8 & average_auc<0.9)

cos.sim.sbs4 <- cos.sim.auc.data %>%
  filter(signature=="SBS4")

```

```{r, include=FALSE}
ggplot(cos.sim.auc.data, aes(x = cosine_similarity, y = average_auc))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs( x = "Cosine similarity between baseline COSMIC and FFPE signatures", y = "Average AUC")+
  geom_point(data = cos.sim.auc.outliers, color = "red")+
  geom_point(data = cos.sim.auc.outliers2, color = "purple")+
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
  geom_hline(yintercept=0.9, linetype="dashed", color = "purple")+
  geom_label_repel(data = cos.sim.auc.outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
  geom_label_repel(data = cos.sim.auc.outliers2, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
  geom_label_repel(data = cos.sim.sbs4, aes(label = signature), max.overlaps = 10, label.size =   0.1)

#ggsave("Fig2b.png", dpi = 300, width = 175, height = 108, units = "mm")

```

# Scatterplot of Average Sensitivity and Specificity vs. Cosine Similarity to FFPE

```{r, include=FALSE}
set.seed(500)
#outer for loop: loop through each COSMIC v3 signature
average.sensitivity.values <- c()
average.specificity.values <- c()
for (x in 1:length(signature.names)){
  signature.sensitivity.vec <- c()
  signature.specificity.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df)
    sig.name <- class.df.colnames[3]
    #Count number of true positives (for sensititivity)
    sensitivity.df <- classification.df %>%
      filter(truth == sig.name & classify == sig.name)
    #Calculate sensitivity
    sensitivity.val = nrow(sensitivity.df)/100
    #Add to the vector of 10 sensitivity values
    signature.sensitivity.vec[y] = sensitivity.val
    #Count number of true negatives (for specificity)
    specificity.df <- classification.df %>%
      filter(truth == "FFPE" & classify == "FFPE")
    #Calculate specificity
    specificity.val = nrow(specificity.df)/100
    #Add to the vector of 10 specificity values
    signature.specificity.vec[y] = specificity.val
  }
  average.signature.sensitivity <- mean(signature.sensitivity.vec)
  average.sensitivity.values[x] <- average.signature.sensitivity
  average.signature.specificity <- mean(signature.specificity.vec)
  average.specificity.values[x] <- average.signature.specificity
}
```

```{r, include=FALSE}
sens.spec.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.sensitivity.values, average.specificity.values)

colnames(sens.spec.data) <- c("signature", "cosine_similarity", "sensitivity", "specificity")

ggplot(sens.spec.data, aes(x=sensitivity)) +
  geom_histogram(color="black", fill="lightblue")+
  labs( x = "Average sensitivity", y = "Count")
#ggsave("supp.fig.3.png", dpi = 300, width = 175, height = 108, units = "mm")

median(sens.spec.data$sensitivity)
IQR(sens.spec.data$sensitivity)

outliers <- sens.spec.data %>%
  filter(sensitivity<0.8 | signature=="SBS4" | signature=="SBS6")

ref.signatures <- sens.spec.data %>%
  filter(signature=="SBS4" )

ggplot(sens.spec.data, aes(x = cosine_similarity, y = sensitivity))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs( x = "Cosine similarity between baseline COSMIC and FFPE signatures", y = "Average sensitivity")+
  geom_point(data = outliers, color = "red")+
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
  geom_label_repel(data = outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)

#ggsave("supp.fig6.png", dpi = 300, width = 175, height = 108, units = "mm")

model.sens <- lm(sensitivity ~ cosine_similarity, data = sens.spec.data)

summary(model.sens)

ggplot(sens.spec.data, aes(x=specificity)) +
  geom_histogram(color="black", fill="lightblue")+
  labs( x = "Average specificity", y = "Count")
#ggsave("supp.fig.4.png", dpi = 300, width = 175, height = 108, units = "mm")

median(sens.spec.data$specificity)
IQR(sens.spec.data$specificity)

outliers <- sens.spec.data %>%
  filter(specificity<0.8 | signature=="SBS4" | signature=="SBS6")

ggplot(sens.spec.data, aes(x = cosine_similarity, y = specificity))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs( x = "Cosine similarity between baseline COSMIC and FFPE signatures", y = "Average specificity")+
  geom_point(data = outliers, color = "red")+
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
  geom_label_repel(data = outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)

#ggsave("supp.fig7.png", dpi = 300, width = 175, height = 108, units = "mm")

model.spec <- lm(specificity ~ cosine_similarity, data = sens.spec.data)

summary(model.spec)

```

# Dilution performance characteristics

```{r}
cosmic.sig4 = signature.matrix.list[[4]]
cosmic.sig6 = signature.matrix.list[[6]]
ffpe.sig = ffpe.signature
```

## Dilution predictions

```{r}
num.mut = 1000
step = seq(10, 990, by = 10)
```

```{r}
probabilities.MMR <- vector(mode = "list", length = length(step))
truth.MMR <- vector(mode = "list", length = length (step))

for (x in 1:length(step)){
  sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x])
  ffpe.truth <- rep("FFPE", step[x])
  sample.sig6 <- create_signature_sample_vector(cosmic.sig6, num.mut-step[x])
  sig6.truth <- rep("MMR", num.mut-step[x])
  test <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"),
                                list(cosmic.sig6, ffpe.sig))
  classification <- test %>%
    mutate(trueBinary = ifelse(truth == "MMR",1,0))

  probabilities.MMR[[x]] <- classification$MMR
  truth.MMR[[x]] <- classification$trueBinary
}
```

```{r}
probabilities.tob <- vector(mode = "list", length = length(step))
truth.tob <- vector(mode = "list", length = length(step))

for (x in 1:length(step)){
  sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x])
  ffpe.truth <- rep("FFPE", step[x])
  sample.sig4 <- create_signature_sample_vector(cosmic.sig4, num.mut-step[x])
  sig4.truth <- rep("Tobacco", num.mut-step[x])
  test <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"),
                                list(cosmic.sig4, ffpe.sig))
  classification <- test %>%
    mutate(trueBinary = ifelse(truth == "Tobacco",1,0))

  probabilities.tob[[x]] <- classification$Tobacco
  truth.tob[[x]] <- classification$trueBinary
}
```

```{r}
test.MMR <- prediction(probabilities.MMR, truth.MMR)
test.tob <- prediction(probabilities.tob, truth.tob)
```

```{r}
evaluate_perf <- function(num_iter, length, pred_obj, perf_method){
  AVG.sens <- vector(mode = "list", length = length)

  for (iteration in 1:num_iter){
    sens.obj <- performance(pred_obj, measure = perf_method)

    for(ind in 1:length(sens.obj@x.values)){
      thresh = as.vector(slot(sens.obj, "x.values")[[ind]])
      sens = as.vector(slot(sens.obj, "y.values")[[ind]])
      if (length(thresh) > 2){
        if (identical(which(thresh<=0.5), integer(0))){
          AVG.sens[[ind]][iteration] = NA
        }
        else{
          low.ind = max(which(thresh >=0.5))
          high.ind = min(which(thresh <=0.5))
          #point slop formula of a line
          slope = (sens[high.ind] - sens[low.ind]) / (thresh[high.ind] - thresh[low.ind])
          AVG.sens[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + sens[low.ind]
        }
      }
      else{
        AVG.sens[[ind]][iteration] = NA
          }
        }
    }

  sens.vec = c()
  for (avg in 1:length(AVG.sens)){
    sens.vec[avg] = mean(AVG.sens[[avg]], na.rm = TRUE)
  }

  return(sens.vec)
}
```

## Sensitivity

```{r}
sens.MMR.vec = evaluate_perf(100, length(step), test.MMR, "sens")
sens.tob.vec = evaluate_perf(100, length(step), test.tob, "sens")
```

### Sensitivity plot

```{r}
sensitivity = append(sens.MMR.vec, sens.tob.vec)
signature.sens = append(rep("SBS6", length(sens.MMR.vec)),
                        rep("SBS4", length(sens.tob.vec)))
dilution = append(step/10, step/10)
sens.df <- data.frame(dilution, sensitivity, signature.sens)

MMR.tob.sens <- ggplot(data = sens.df, mapping = aes(x = dilution, 
                                     y = sensitivity, 
                                     linetype = signature.sens)) +
  geom_point(aes(shape = signature.sens)) +
  geom_smooth(se=FALSE) +
  labs(x = "Percent of FFPE signature",
       y = "Sensitivity",
       linetype = "Baseline signature",
       shape = "Baseline signature") +
  coord_cartesian(xlim = c(0,100), ylim = c(0,1))+
  scale_y_continuous(breaks=seq(from=0, to=1, by=.2))+
  geom_hline(yintercept=0.8, linetype=4, color = "red")

#ggsave("fig3a.png", plot = MMR.tob.sens, dpi = 300, width = 175, height = 108, units = "mm")

MMR.tob.sens

```

### Sensitivity threshold for all signatures

```{r}
#set threshold for a decent sensitivity for our classifier
optimal.sens = 0.8
max.sens.noise = c()
#loop through each signature
for (sig in 1:length(signature.matrix.list)){
  #empty lists to hold classifier outputs and ground truths
  probabilities.sig <- vector(mode = "list", length = length(step))
  truth.sig <- vector(mode = "list", length = length(step))

  #dilute each signature with some percent of FFPE
  for (x in 1:length(step)){
    sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x])
    ffpe.truth <- rep("FFPE", step[x])
    sample.sig <- create_signature_sample_vector(signature.matrix.list[[sig]],
                                                 num.mut-step[x])
    sig.truth <- rep("Biological", num.mut-step[x])
    #run classifier
    test <- get_classification_df(list(sample.sig, sample.ffpe), c("Biological","FFPE"),
                                  list(signature.matrix.list[[sig]], ffpe.sig))
    classification <- test %>%
      mutate(trueBinary = ifelse(truth == "Biological",1,0))
    #extract classifications and ground truth
    probabilities.sig[[x]] <- classification$Biological
    truth.sig[[x]] <- classification$trueBinary
  }

  #create ROC prediction object and evaluate
  test.sig <- prediction(probabilities.sig, truth.sig)
  sens.sig.vec = evaluate_perf(100, length(step), test.sig, "sens")

  #cut off noise past our threshold
  if(identical(which(sens.sig.vec <= optimal.sens), integer(0))){
    max.sens.noise[sig] = 100
    }
  else{
    max.sens.noise[sig] = min(which(sens.sig.vec <= optimal.sens))
    }
}
```

```{r}
#visualize results

all.sens.df <- data.frame(signature.names, max.sens.noise, ffpe.cosmic.cos.sims)
cos.sim.sbs4.sbs6 <- all.sens.df %>%
  filter(signature.names %in% c("SBS4", "SBS6"))
cos.sim.sens.outliers <- all.sens.df %>%
  filter(max.sens.noise <= 50)

all.sens <- ggplot(data = all.sens.df, mapping = aes(x = ffpe.cosmic.cos.sims,
                                         y = max.sens.noise))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  labs(x = "Cosine Similarity to FFPE",
       y = "Maximum Percent of FFPE Noise") +
  coord_cartesian(ylim = c(0,100), xlim = c(0,1)) +
  geom_hline(yintercept=50, linetype="dashed", color = "red")+
  geom_label_repel(data = cos.sim.sbs4.sbs6, aes(label = signature.names), 
                   max.overlaps = 10,
                   label.size =   0.1)+
  geom_label_repel(data = cos.sim.sens.outliers, aes(label = signature.names), 
                   max.overlaps = 10,
                   label.size =   0.1)

#ggsave("fig3c_sens.png", plot = all.sens, dpi = 300, width = 175, height = 108, units = "mm")
all.sens
```

```{r}
#linear relationship between cosine similarity and max FFPE noise
model.all.sens <- lm(max.sens.noise ~ ffpe.cosmic.cos.sims, data = all.sens.df)
summary(model.all.sens)
```

## Specificity

```{r}
spec.MMR.vec = evaluate_perf(100, length(step), test.MMR, "spec")
spec.tob.vec = evaluate_perf(100, length(step), test.tob, "spec")
```

### Specificity plot

```{r}
specificity = append(spec.MMR.vec, spec.tob.vec)
signature.spec = append(rep("SBS6", length(spec.MMR.vec)),
                        rep("SBS4", length(spec.tob.vec)))
dilution = append(step/10, step/10)
spec.df <- data.frame(dilution, specificity, signature.spec)

MMR.tob.spec <- ggplot(data = spec.df, mapping = aes(x = dilution, 
                                     y = specificity, 
                                     lty = signature.spec)) +
  geom_point(aes(shape = signature.spec)) +
  geom_smooth() +
  labs(x = "Percent of FFPE signature",
       y = "Specificity",
       lty = "Baseline signature",
       shape = "Baseline signature") +
  coord_cartesian(xlim = c(0,100), ylim = c(0,1))+
    scale_y_continuous(breaks=seq(from=0, to=1, by=.2))+
  geom_hline(yintercept=0.8, linetype=4, color = "red")

#ggsave("fig3b.png", plot = MMR.tob.spec, dpi = 300, width = 175, height = 108, units = "mm")

MMR.tob.spec
```

### Specificity threshold for all signatures

```{r}
#set threshold for a decent specificity for our classifier
optimal.spec = 0.8
min.spec.noise = c()
#loop through each signature
for (sig in 1:length(signature.matrix.list)){
  #empty lists to hold classifier outputs and ground truths
  probabilities.sig <- vector(mode = "list", length = length(step))
  truth.sig <- vector(mode = "list", length = length(step))

  #dilute each signature with some percent of FFPE
  for (x in 1:length(step)){
    sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x])
    ffpe.truth <- rep("FFPE", step[x])
    sample.sig <- create_signature_sample_vector(signature.matrix.list[[sig]],
                                                 num.mut-step[x])
    sig.truth <- rep("Biological", num.mut-step[x])
    #run classifier
    test <- get_classification_df(list(sample.sig, sample.ffpe), c("Biological","FFPE"),
                                  list(signature.matrix.list[[sig]], ffpe.sig))
    classification <- test %>%
      mutate(trueBinary = ifelse(truth == "Biological",1,0))
    #extract classifications and ground truth
    probabilities.sig[[x]] <- classification$Biological
    truth.sig[[x]] <- classification$trueBinary
  }

  #create ROC prediction object and evaluate
  test.sig <- prediction(probabilities.sig, truth.sig)
  spec.sig.vec = evaluate_perf(100, length(step), test.sig, "spec")

  #cut off noise past our threshold
  if(identical(which(spec.sig.vec <= optimal.spec), integer(0))){
    min.spec.noise[sig] = 1
    }
  else{
    min.spec.noise[sig] = max(which(spec.sig.vec <= optimal.spec))
    }
}
```

```{r}
#visualize results

all.spec.df <- data.frame(signature.names, min.spec.noise, ffpe.cosmic.cos.sims)
cos.sim.sbs4.sbs6 <- all.spec.df %>%
  filter(signature.names == "SBS4")
cos.sim.spec.outliers <- all.spec.df %>%
  filter(min.spec.noise >= 50)

all.spec <- ggplot(data = all.spec.df, mapping = aes(x = ffpe.cosmic.cos.sims,
                                         y = min.spec.noise))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  labs(x = "Cosine Similarity to FFPE",
       y = "Minimum Percent of FFPE Noise") +
  coord_cartesian(ylim = c(0,100), xlim = c(0,1)) +
  geom_hline(yintercept=50, linetype="dashed", color = "red")+
  geom_label_repel(data = cos.sim.sbs4.sbs6, aes(label = signature.names), 
                   max.overlaps = 10,
                   label.size =   0.1)+
  geom_label_repel(data = cos.sim.spec.outliers, aes(label = signature.names), 
                   max.overlaps = 10,
                   label.size =   0.1)

#ggsave("fig3d_spec.png", plot = all.spec, dpi = 300, width = 175, height = 108, units = "mm")
all.spec
```

```{r}
model.all.spec <- lm(min.spec.noise ~ ffpe.cosmic.cos.sims, data = all.spec.df)
summary(model.all.spec)
```

## Reconstructed profiles of diluted MMR

```{r, include=FALSE}
dilution.sig = signature.6.mat
dilution.sig.name = "SBS6"
total.mut = 2000
dilute = c(0.1,0.3,0.5,0.7,0.9)*total.mut
dilution.list <- vector(mode = "list", length = 5)
i = 1

for (mut in dilute){
  #create samples with some percent of noise
  add.sig = create_signature_sample_vector(dilution.sig, total.mut-mut)
  sig.profile = vector_to_matrix(add.sig)
  ffpe.sample = create_signature_sample_vector(ffpe.signature, mut)
  ffpe.sig = vector_to_matrix(c(ffpe.sample, add.sig))
  #classify mutations
  dilution.df <- get_classification_df(list(add.sig, ffpe.sample), c("SBS6", "FFPE"),
                                       list(dilution.sig, ffpe.signature))
  #filter FFPE mutations
  dilution.df.filtered <- dilution.df %>%
      filter(classify != "FFPE")
  recon.sig <- vector_to_matrix(dilution.df.filtered$mutations)

  #values for performance characteristics
  performance.df <- dilution.df %>%
    mutate(TP = ifelse(classify == dilution.sig.name & truth == dilution.sig.name, 1,0),
           FP = ifelse(classify == dilution.sig.name & truth == "FFPE", 1,0),
           TN = ifelse(classify == "FFPE" & truth == "FFPE", 1,0),
           FN = ifelse(classify == "FFPE" & truth == dilution.sig.name, 1,0))
  TN = performance.df$TN
  FN = performance.df$FN
  TP = performance.df$TP
  FP = performance.df$FP
  acc.vec = TP + TN

  #calculate performance characterisitics
  sens.value = round(mean(TP)/(mean(TP) + mean(FN)), digits = 2)
  spec.value = round(mean(TN)/(mean(TN) + mean(FP)), digits = 2)
  acc.value = round(mean(acc.vec), digits = 2)

  sig.plot <- plot_96_profile(sig.profile)+labs(title = "Signature 6 simulated sample", 
                                                x = "Context")
  ffpe.sig.plot <- plot_96_profile(ffpe.sig)+labs(title = "Simulated Sample with FFPE noise",
                                                  x = "Context")
  recon.plot <- plot_96_profile(recon.sig)+labs(title = "Signature 6 Reconstructed", 
                                                x = "Context", 
                                                caption = paste0("Sensitivity: ", sens.value,
                                                                 " | Specificity: ",
                                                                 spec.value, " | Accuracy: ",
                                                                 acc.value))
  dilution.list[[i]] <- grid.arrange(sig.plot, ffpe.sig.plot, recon.plot)

  i = i+1
}

#ggsave("sig6reconstruction_1.png", plot = dilution.list[[1]], dpi = 300, width = 175, height = 243, units = "mm")
#ggsave("sig6reconstruction_2.png", plot = dilution.list[[2]], dpi = 300, width = 175, height = 243, units = "mm")
#ggsave("sig6reconstruction_3.png", plot = dilution.list[[3]], dpi = 300, width = 175, height = 243, units = "mm")
#ggsave("sig6reconstruction_4.png", plot = dilution.list[[4]], dpi = 300, width = 175, height = 243, units = "mm")
#ggsave("sig6reconstruction_5.png", plot = dilution.list[[5]], dpi = 300, width = 175, height = 243, units = "mm")
```

## File of all signatures diluted with FFPE noise

```{r, include=FALSE}
#columns in data frame
sens = vector(mode = "list", length = length(signature.names))
spec = vector(mode = "list", length = length(signature.names))
acc = vector(mode = "list", length = length(signature.names))
signature = vector(mode = "list", length = length(signature.names))

#create data frame values 10 times
iteration.num = 3
for (iteration in 1:iteration.num){
  #iterate through each signature
  for (sig.num in 1:length(signature.names)){
    #store name of signature (once for every dilution)
    signature[[sig.num]] = append(signature[[sig.num]],
                                  rep(signature.names[sig.num], 5))
    #store values generated by get_classification_df for each dilution
    probabilities <- vector(mode = "list", length = 5)
    truth <- vector(mode = "list", length = 5)
    #initialize accumulator
    i = 1
      #each signature is diluted with 10%, 30%, 50%, 70%, and 90% FFPE noise
      # (1000 mutations total each time)
      for (x in c(100, 300, 500, 700, 900)){
        #create some percent of FFPE noise
        sample.ffpe <- create_signature_sample_vector(ffpe.signature, x)
        ffpe.truth <- rep("FFPE", x)
        #fill the rest with true signal
        sample <- create_signature_sample_vector(signature.matrix.list[[sig.num]], 1000-x)
        sample.truth <- rep(signature.names[sig.num], 1000-x)
        #evaluate using Bayes
        test <- get_classification_df(list(sample, sample.ffpe), 
                                      c(signature.names[sig.num],"FFPE"),
                                      list(signature.matrix.list[[sig.num]],
                                           ffpe.signature))
        #truth column turned into binary for ROCR package
        classification <- test %>%
          mutate(trueBinary = ifelse(truth == signature.names[sig.num],1,0))
        #store probabilities and truths for particular dilution
        probabilities[[i]] <- classification[[3]]
        truth[[i]] <- classification$trueBinary
        i = i + 1
      }
    #evaluate all 5 dilution samples
    pred = prediction(probabilities, truth)
    acc[[sig.num]] = append(acc[[sig.num]], evaluate_perf(50, 5, pred, "acc"))
    sens[[sig.num]] = append(sens[[sig.num]], evaluate_perf(50, 5, pred, "sens"))
    spec[[sig.num]] = append(spec[[sig.num]], evaluate_perf(50, 5, pred, "spec"))
  }
}

#vector indicating dilution amount
percent.FFPE = rep(c("10%", "30%", "50%", "70%", "90%"),
                   length(signature.names)*iteration.num)
#vectorize lists
signature = unlist(signature)
acc = unlist(acc)
sens = unlist(sens)
spec = unlist(spec)

#convert into data frame
table = data.frame(percent.FFPE, signature, round(sens, digits = 2), 
                   round(spec, digits = 2), round(acc, digits = 2))

#avg the multiple sensitivity/specificity values for each dilution and each signature
table <- table %>%
  group_by(percent.FFPE, signature) %>%
  summarize(sens = mean(sens, na.rm = TRUE),
            spec = mean(spec, na.rm = TRUE),
            acc = mean(acc, na.rm = TRUE)) %>%
  mutate(signature = factor(signature, signature.names)) %>%
  arrange(signature)
```

<!-- # Mutational load performance characteristics -->

<!-- ## Mutational load predictions -->

<!-- ```{r} -->

<!-- step_load = seq(50, 1000, by = 50) -->

<!-- MMR.probabilities <- vector(mode = "list", length = length(step)) -->

<!-- MMR.truth <- vector(mode = "list", length = length(step)) -->

<!-- for (x in 1:length(step)){ -->

<!--   sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x]/2) -->

<!--   ffpe.truth <- rep("FFPE", step[x]/2) -->

<!--   sample.sig6 <- create_signature_sample_vector(cosmic.sig6, step[x]/2) -->

<!--   sig6.truth <- rep("MMR", step[x]/2) -->

<!--   test <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"), -->

<!--                                 list(cosmic.sig6, ffpe.sig)) -->

<!--   classification <- test %>% -->

<!--     mutate(trueBinary = ifelse(truth == "MMR",1,0)) -->

<!--   MMR.probabilities[[x]] <- classification$MMR -->

<!--   MMR.truth[[x]] <- classification$trueBinary -->

<!-- } -->

<!-- ``` -->

<!-- ```{r} -->

<!-- tob.probabilities <- vector(mode = "list", length = length(step)) -->

<!-- tob.truth <- vector(mode = "list", length = length(step)) -->

<!-- for (x in 1:length(step)){ -->

<!--   sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x]/2) -->

<!--   ffpe.truth <- rep("FFPE", step[x]/2) -->

<!--   sample.sig4 <- create_signature_sample_vector(cosmic.sig4, step[x]/2) -->

<!--   sig4.truth <- rep("tob", step[x]/2) -->

<!--   test <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"), -->

<!--                                 list(cosmic.sig4, ffpe.sig)) -->

<!--   classification <- test %>% -->

<!--     mutate(trueBinary = ifelse(truth == "Tobacco",1,0)) -->

<!--   tob.probabilities[[x]] <- classification$Tobacco -->

<!--   tob.truth[[x]] <- classification$trueBinary -->

<!-- } -->

<!-- ``` -->

<!-- ```{r} -->

<!-- load.MMR <- prediction(MMR.probabilities, MMR.truth) -->

<!-- load.tob <- prediction(tob.probabilities, tob.truth) -->

<!-- ``` -->

<!-- ## Sensitivity -->

<!-- ```{r} -->

<!-- load.sens.MMR.vec = evaluate_perf(100, length(step_load), load.MMR, "sens") -->

<!-- load.sens.tob.vec = evaluate_perf(100, length(step_load), load.tob, "sens") -->

<!-- ``` -->

<!-- ```{r} -->

<!-- #plot(step, load.sens.MMR.vec, ylim = c(0,1)) -->

<!-- sensitivity = append(load.sens.MMR.vec, load.sens.tob.vec) -->

<!-- signature.sens = append(rep("MMR", length(load.sens.MMR.vec)), -->

<!--                         rep("Tobacco", length(load.sens.tob.vec))) -->

<!-- dilution = append(step, step) -->

<!-- sens.df <- data.frame(dilution, sensitivity, signature.sens) -->

<!-- ggplot(data = sens.df, mapping = aes(x = dilution,  -->

<!--                                      y = sensitivity,  -->

<!--                                      color = signature.sens)) + -->

<!--   geom_point() + -->

<!--   geom_smooth() + -->

<!--   labs(title = "Sensitivity as a function of mutational load", -->

<!--        x = "Number of mutations in each sample", -->

<!--        y = "Sensitivity value", -->

<!--        color = "True signal") + -->

<!--   coord_cartesian(ylim = c(0,1)) -->

<!-- ``` -->

<!-- ## Specificity -->

<!-- ```{r} -->

<!-- load.spec.MMR.vec = evaluate_perf(100, length(step.load), load.MMR, "spec") -->

<!-- load.spec.tob.vec = evaluate_perf(100, length(step.load), load.tob, "spec") -->

<!-- ``` -->

<!-- ```{r} -->

<!-- #plot(step, load.spec.MMR.vec, ylim = c(0,1)) -->

<!-- specificity = append(load.spec.MMR.vec, load.spec.tob.vec) -->

<!-- signature.spec = append(rep("MMR", length(load.spec.MMR.vec)), -->

<!--                         rep("Tobacco", length(load.spec.tob.vec))) -->

<!-- dilution = append(step, step) -->

<!-- spec.df <- data.frame(dilution, specificity, signature.spec) -->

<!-- ggplot(data = spec.df, mapping = aes(x = dilution,  -->

<!--                                      y = specificity,  -->

<!--                                      color = signature.spec)) + -->

<!--   geom_point() + -->

<!--   geom_smooth() + -->

<!--   labs(title = "Specificity as a function of mutational load", -->

<!--        x = "Number of mutations in each sample", -->

<!--        y = "Specificity value", -->

<!--        color = "True signal") + -->

<!--   coord_cartesian(ylim = c(0,1)) -->

<!-- ``` -->