### 2022-08-03 paper_analysis.Rmd
Report and figures for the paper

---
title: "Excerno analysis"
author: "Audrey Mitchell, Marco Ruiz, Soua Yang, Jaime Davila"
date: "8/3/2022"
output:
  html_document: 
    keep_md: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache = TRUE, 
                      warning=FALSE, message = FALSE)
```

# Set up

We start our analysis by loading our libraries and functions from other files

```{r}
library(NMF)
library(MutationalPatterns)
library(tidyverse)
library(ggrepel)
library(ROCR)
library(kableExtra)
library(gridExtra)
library(broom)
library(plotROC)
source("bayes_function.R")
source("get_classification_df.R")
source("simulation.R")
```

We also define the function `vector_to_matrix` which allows to switch between representations of mutational profiles, in particular to use `plot_96_profile` from `MutationalPatterns`

```{r}
vector_to_matrix <- function(mutations.vector){
  tab <- table(mutations.vector)/sum(table(mutations.vector))
  #transform table to a data frame and assign column names to mutations and frequencies
  mutations.df <- data.frame(tab)
  colnames(mutations.df) <- c("mutations", "frequencies")

  #defines a function that joins two data frames and replaces NA with 0
  left_join_NA <- function(x, y, by) {
    left_join(x = x, y = y, by = by) %>% 
      mutate_each(funs(replace(., which(is.na(.)), 0)))
  }

  #converts vector of 96 mutation types to a data frame
  mutation.types.df <- data.frame(mutations)
  #joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
  complete.mutations.df <- left_join_NA(mutation.types.df, mutations.df, by = "mutations")

  #converts data frame of mutation types and frequencies to a data frame and assigns rownames
  mutations.matrix <- as.matrix(complete.mutations.df$frequencies)
  rownames(mutations.matrix) = mutations

  return(mutations.matrix)
}
```

We start by loading the FFPE mutational signatures and the SBS mutational signatures from COSMIC:

```{r}
#load FFPE signature as a matrix
ffpe_signature <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))

#create vector of the 96 mutation types
mutations <- rownames(ffpe_signature)

#Load COSMIC signatures (version 3)
cosmic_signatures <- get_known_signatures(muttype = "snv")

#Create vector of the signature names
signature_names <- colnames(cosmic_signatures)
```

And we create a list with all the COSMIC mutational signatures

```{r}
#Create list of signature matrices with assigned rownames
signature_matrix_list <- vector ("list", length(signature_names))
for (i in 1:length(signature_names)){
  sig <- cosmic_signatures[,i]
  sig_mat <- as.matrix(sig)
  rownames(sig_mat) <- mutations
  colnames(sig_mat) <- signature_names[i]
  signature_matrix_list[[i]] <- sig_mat
}
```

# Initial testing

We created two simulated datasets as follows:

* We combine 500 mutations from SBS4 and 500 mutations from FFPE

* We combine 500 mutations from SBS6 and 500 mutations from FFPE

## First dataset: SBS4

Let's create and visualize our first dataset by simulating mutations from SBS4

```{r}
set.seed(500)
signature_4_mat <- as.matrix(cosmic_signatures[,"SBS4"])
rownames(signature_4_mat) <- mutations

sig4_500 <- create_signature_sample_vector(signature_4_mat, 500)
sig4_profile <- vector_to_matrix(sig4_500)

plot_96_profile(sig4_profile)+
  labs(title = "Signature 4 Simulated Sample", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))
```

Then, let's add the FFPE noise and visualize the resulting mutational profile

```{r}
addffpe_500 <- create_signature_sample_vector(ffpe_signature, 500)
sig4_ffpe <- vector_to_matrix(c(sig4_500, addffpe_500))

plot_96_profile(sig4_ffpe)+
  labs(title = "Simulated Sample with FFPE Noise", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))
```

Now, let's apply our method on the simulated dataset:

```{r}
sig4_ffpe_tbl <- get_classification_df(
  list(sig4_500, addffpe_500),
  c("SBS4", "FFPE"), 
  list(signature_4_mat, ffpe_signature)) %>%
  as_tibble()
```

And calculate the confusion matrix

```{r}
(conf_mat <- table(sig4_ffpe_tbl$truth, sig4_ffpe_tbl$classify))
```

Our method is very good at identifying mutations that originated from the FFPE process, although some mutations originated from SBS4 gets missclassified as FFPE. Let's calculate the performance metrics:

```{r}
(sig4_sensitivity <- (conf_mat[2,2] / (conf_mat[2,2]+ conf_mat[2,1])))
(sig4_specificity <- (conf_mat[1,1] / (conf_mat[1,1]+ conf_mat[1,2])))
```

## Second dataset: SBS6

Let's repeat the same experiment, this time using SBS6 as our baseline signature

```{r}
signature_6_mat <- as.matrix(cosmic_signatures[,"SBS6"])
rownames(signature_6_mat) <- mutations
sig6_500 <- create_signature_sample_vector(signature_6_mat, 500)

sig6_ffpe_tbl <- get_classification_df(
  list(sig6_500, addffpe_500),
  c("SBS6", "FFPE"),
  list(signature_6_mat, ffpe_signature))%>%
  as_tibble()
```

And let's calculate our confusion matrix and our performance characteristics

```{r}
(conf_mat <- table(sig6_ffpe_tbl$truth, sig6_ffpe_tbl$classify))
(sig6_sensitivity <- (conf_mat[2,2] / (conf_mat[2,2]+ conf_mat[2,1])))
(sig6_specificity <- (conf_mat[1,1] / (conf_mat[1,1]+ conf_mat[1,2])))
```

It seems this time our method did not perform as well. Let's create mutational profile plots for:

* The original profile (SBS6)
* The original profile (SBS6) plus the FFPE noise
* The profile obtained after filtering FFPE using `excerno`

```{r}
# Original profile
sig6_profile <- vector_to_matrix(sig6_500)
sig6_profile_plot <- plot_96_profile(sig6_profile)+
  labs(title = "Signature 6 Simulated Sample", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))
# Profile + FFPE noise
sig6_ffpe <- vector_to_matrix(c(sig6_500, addffpe_500))
sig6_ffpe_plot <- plot_96_profile(sig6_ffpe)+
  labs(title = "Simulated Sample with FFPE Noise", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))
# Profile after excerno
sig6_ffpe_filtered <- sig6_ffpe_tbl %>%
  filter(classify != "FFPE")
sig6_recon_matrix <- vector_to_matrix(sig6_ffpe_filtered$mutations)

sig6_recon_plot <- plot_96_profile(sig6_recon_matrix)+
  labs(title = "Signature 6 Reconstructed", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

(sig6_grid <- grid.arrange(sig6_profile_plot, sig6_ffpe_plot, sig6_recon_plot))
ggsave("paper_figures/sup1.png", sig6_grid, width = 175, height = 243, units = "mm")

```

It seems filtered SBS6 dataset differs from the original signature, particularly in C>T mutations preceded by an A or a T

## ROC curve for both datasets

We would like to combine the information from both datasets (SBS4 and SBS6) in order to plot them together.

```{r}
sig4_tbl <- sig4_ffpe_tbl %>%
  select(truth, SBS4) %>%
  rename (bayes = SBS4) %>%
  mutate (dataset = "SBS4",
          truth = if_else (truth=="SBS4",1, 0))

sig6_tbl <- sig6_ffpe_tbl %>%
  select(truth, SBS6) %>%
  rename (bayes = SBS6) %>%
  mutate (dataset = "SBS6",
          truth = if_else (truth=="SBS6",1, 0))

allsim_tbl <- rbind (sig4_tbl, sig6_tbl)
```

And we generate the ROC curve for both datasets and calculate their AUC

```{r}
(roc_plot <- 
  ggplot(allsim_tbl, aes(d=truth, m=bayes, color=dataset))+
  geom_roc(cutoffs.at = c(0.5), cutoff.labels = c(0.5)) +
  labs(x="False Positive Rate",
       y="True Positive Rate",
       color="Signature"))

ggsave("paper_figures/fig2a.png", roc_plot, width = 131.25, height = 108, units = "mm")

calc_auc(roc_plot)
```

# When is `excerno` a good classifier?

We hypothesize that the performance of our classifier depends on how different the base mutational signature is from the FFPE signature.

## Initial Setup

A straightforward way of calculating the similarity across signatures the cosine similarity. So we first generate the cosine similarity of all the COSMIC signatures against the FFPE signature:

```{r}
ffpe_cosmic_cos_sims <- vector(length=length(signature_names))
for (i in 1:length(signature_names)){
  cos_sim <- cos_sim_matrix(signature_matrix_list[[i]], ffpe_signature)
  ffpe_cosmic_cos_sims[i] <- cos_sim[1,1]
}
```

Second, we will create a function that mixes mutations from FFPE and another given mutational signature and calls `excerno`

```{r}
test_mix_mutations <- function (cosmic_idx, num_cosmic, num_ffpe) {

  # Create baseline mutations
  baseline_sig <- signature_matrix_list[[cosmic_idx]]
  baseline_name <- signature_names[cosmic_idx]
  baseline_mut_vec <- create_signature_sample_vector(baseline_sig, num_cosmic)

  # Create FFPE mutations
  ffpe_mut_vec <- create_signature_sample_vector(ffpe_signature, num_ffpe)

  # Call excerno classification and format as a tibble
  get_classification_df (list(baseline_mut_vec, ffpe_mut_vec), 
                          c(baseline_name,"FFPE"), 
                          list(baseline_sig, ffpe_signature)) %>% 
    as_tibble()
}
```

And we try `test_mix_mutations` and plot the results to check it is working:

```{r}
set.seed(12345)
test_tbl <- test_mix_mutations(1,500,500)

test_tbl %>%
  select (mutations, truth, SBS1, FFPE) %>%
  pivot_longer(c(SBS1, FFPE)) %>%
  ggplot(aes(x=truth, y=value, fill=name)) +
    geom_boxplot()

```

Now let's create a function that calculates the performance metrics (sensitivity, specificity, accuracy) on the output of `test_mix_mutations`

```{r}
calc_metrics_mix <- function (cosmic_idx, num_cosmic, num_ffpe) {

  baseline_name <- signature_names[cosmic_idx]
  test_tbl <- test_mix_mutations(cosmic_idx,num_cosmic,num_ffpe)

  tp <- test_tbl %>%
    filter(truth == baseline_name & classify == baseline_name) %>%
    count() %>%
    pull(n)

  fp <- test_tbl %>%
    filter(truth != baseline_name & classify == baseline_name) %>%
    count() %>%
    pull(n)

  fn <- test_tbl %>%
    filter(truth == baseline_name & classify == "FFPE") %>%
    count() %>%
    pull(n)

  tn <- test_tbl %>%
    filter(truth == "FFPE" & classify == "FFPE") %>%
    count() %>%
    pull(n)

  tibble(
    Sensitivity = tp / (tp +fn),
    Specificity = tn / (tp + fn),
    Accuracy = (tp +tn) /(tp+tn+fp+fn))
}

```

And we test our method:

```{r}
set.seed(12345)

calc_metrics_mix(1,500,500)
calc_metrics_mix(4,500,500)
calc_metrics_mix(6,500,500)
```

We create a function that calculates the average metrics over a fixed number of simulations `num_iter` and test it

```{r}
calc_metrics_mix_num <- function (cosmic_idx, num_iter, num_cosmic, num_ffpe) {

  baseline_name <- signature_names[cosmic_idx]

  1:num_iter %>% 
    map_dfr(~calc_metrics_mix(cosmic_idx,num_cosmic,num_ffpe))  %>%
    pivot_longer(1:3) %>% 
    group_by(name) %>%
    summarize(mean = mean(value)) %>%
    mutate(sbs = baseline_name)
}

calc_metrics_mix_num(1,10,500,500)
calc_metrics_mix_num(4,10,500,500)
calc_metrics_mix_num(6,10,500,500)
```

## Simulations across all COSMIC signatures

We are ready to generate the data over all of COSMIC mutational signatures, using 10 iterations and 500 mutations for FFPE and 500 mutations from the baseline COSMIC signature

```{r}
sbs_num = length(signature_names)
base_num = 500
ffpe_num = 500
reps = 10

set.seed(123456)
sim_tbl <- 1:sbs_num %>%
  map_dfr(~calc_metrics_mix_num(.x, reps, base_num, ffpe_num))
```

First we summarize the metrics across all the COSMIC mutational signatures

```{r}
sim_tbl <- sim_tbl %>%
  mutate (name = factor (name, 
                         levels = c("Sensitivity", "Specificity","Accuracy"))) 
sim_tbl %>%
  ggplot(aes(mean))+
  geom_histogram(binwidth = 0.025)+
  facet_grid(.~name) +
  labs(x="Metric across all COSMIC signatures")

ggsave("paper_figures/supp2.png", dpi = 300, width = 175, height = 75, units = "mm")

summary_tbl <- sim_tbl %>%
  group_by(name) %>%
  summarize (median = median (mean),
           IQR = IQR(mean)) %>%
  mutate_if(is.numeric, format, digits=2)

summary_tbl %>%
  kbl() %>%
  kable_styling()
```

## Performance characteristics vs similarity against FFPE signature

Now we explore the relationship of the performance metrics against the similarity between the baseline signature and FFPE using a figure

```{r}
sim_tbl <- sim_tbl %>%
  mutate (ffpe_cos = rep(ffpe_cosmic_cos_sims, each=3))

sim_tbl %>%
  ggplot(aes(x=ffpe_cos, mean)) +
  geom_point()+
  geom_smooth(method="lm", formula = "y~x")+
  facet_grid(~name)
```

We define a function to represent the mutational signatures that fall below a threshold for a particular metric.

```{r}
plot_metrics_excerno <- function (tbl, metric, threshold) {

  # Subset the table to the specific metric  
  new_tbl <- filter(tbl, name==metric) 
  # Identify signature that fall below the threshold
  outliers <- filter( new_tbl, (mean < threshold) | sbs == "SBS4" | sbs == "SBS6")

  ggplot(new_tbl, aes(x = ffpe_cos, y = mean))+
    geom_point()+
    geom_smooth(method = "lm", se=FALSE)+
    geom_point(data = outliers, color = "red")+
    geom_hline(yintercept=threshold, linetype="dashed", color = "red")+
    geom_label_repel(data=outliers, 
                   aes(label = sbs),
                   max.overlaps = 10, 
                   label.size = .1)+
    labs(x = "Cosine similarity against FFPE signature", 
         y = metric )
}
```

And we plot all the metrics using a threshold of 80%

```{r}
plot_metrics_excerno(sim_tbl, "Accuracy", 0.8)
ggsave("paper_figures/fig2b.png", dpi = 300, width = 175, height = 108, units = "mm")

plot_metrics_excerno(sim_tbl, "Sensitivity", 0.8)
ggsave("paper_figures/supp3.png", dpi = 300, width = 175, height = 108, units = "mm")

plot_metrics_excerno(sim_tbl, "Specificity", 0.8)
ggsave("paper_figures/supp4.png", dpi = 300, width = 175, height = 108, units = "mm")
```

And we create auxiliar function for getting the coefficients of the linear model and their fit

```{r}
print_lm_coefs <- function (tbl, metric) {
  new_tbl <-  filter (tbl, name==metric)
  lm_metrics<- lm(mean ~ ffpe_cos, data = new_tbl)

  tidy(lm_metrics) %>%
    mutate_if(is.numeric, format, digits=2)%>%
    kbl() %>%
    kable_styling()
}

print_lm_fit <- function (tbl, metric) {
  new_tbl <-  filter (tbl, name==metric)
  lm_metrics<- lm(mean ~ ffpe_cos, data = new_tbl)

  glance(lm_metrics) %>%
    mutate_if(is.numeric, format, digits=2)%>%
    kbl() %>%
    kable_styling()
}
```

And print the information of the linear model for

* Accuracy

```{r}
print_lm_coefs(sim_tbl, "Accuracy")
print_lm_fit(sim_tbl, "Accuracy")
```

* Sensitivity

```{r}
print_lm_coefs(sim_tbl, "Sensitivity")
print_lm_fit(sim_tbl, "Sensitivity")
```

* Specificity
```{r}
print_lm_coefs(sim_tbl, "Specificity")
print_lm_fit(sim_tbl, "Specificity")
```

```{r}
sessionInfo()
```