### 2022-08-10 sandbox/Bayes Mutation Example.Rmd

---
title: "Bayes Question"
author: "Audrey Mitchell"
date: "6/7/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MutationalPatterns)
library(BSgenome)
library(BiocManager)
install("BSgenome.Hsapiens.UCSC.hg19")
ref_genome <- "BSgenome.Hsapiens.UCSC.hg19"
library(ref_genome, character.only = TRUE)
vcf_files <- list.files(system.file("extdata", package="MutationalPatterns"), pattern = ".vcf", full.names = TRUE)
vcf_files <- vcf_files[-7]
sample_names <- c("blood-AC", "blood-ACC55", "blood-BCH", "colon1", "colon2", "colon3", "intestine1", "intestine2", "intestine3", "liver1", "liver2", "liver3")
vcfs <- read_vcfs_as_granges(vcf_files, sample_names, ref_genome)
tissue <- c(rep("blood", 3), rep("colon",3), rep("intestine",3), rep("liver",3))
type_occurrences <- mut_type_occurrences(vcfs, ref_genome)
mut_mat <- mut_matrix(vcf_list = vcfs, ref_genome = ref_genome)
```

```{r, include = FALSE}
nmf_res <- extract_signatures(mut_mat, rank = 2, nrun = 10)
colnames(nmf_res$signatures) <- c("Signature A", "Signature B")
rownames(nmf_res$contribution) <- c("Signature A", "Signature B")
```

# Given that we observe a A(C>T)A mutation in the blood-AC sample, what is the probability that it came from Signature A?

## Method 1: Splitting Pr(A(C>T)A) into A(C>T)A mutations from Signature A and A(C>T)A mutations from Signature B
$Pr(SigA|A(C>T)A) = \frac{Pr(SigA)Pr(A(C>T)A|SigA)}{Pr(SigA)Pr(A(C>T)A|SigA)+Pr(SigB)Pr(A(C>T)A|SigB)}$

```{r, echo = FALSE}
plot_96_profile(nmf_res$signatures, condensed = TRUE)
```
From the plot above we observe that $Pr(A(C>T)A|SigA)= 0.025$ and $Pr(A(C>T)A|SigB)=0.05$.
```{r, echo= FALSE}
plot_contribution(nmf_res$contribution, nmf_res$signature, mode = "relative", coord_flip = TRUE)
nmf_res$contribution
```

```{r}
0.1093118/(0.2800362+0.1093118)
0.2800362/(0.2800362+0.1093118)
```

From the plot above and the contributions of Signature A and B, we see that for the blood-AC sample $Pr(SigA) \approx 0.28$ and $Pr(SigB) \approx 0.72$.\
Then $Pr(SigA|A(C>T)A) = \frac{Pr(SigA)Pr(A(C>T)A|SigA)}{Pr(SigA)Pr(A(C>T)A|SigA)+Pr(SigB)Pr(A(C>T)A|SigB)}=\frac{0.28(0.025)}{0.28(0.025)+0.72(0.05)}=\frac{0.007}{0.043}=0.1628$

## Method 2: Using Total Pr(A(C>T)A) from the mutational profile of the sample
$Pr(SigA|A(C>T)A)=\frac{Pr(SigA)Pr(A(C>T)A|SigA)}{Pr(A(C>T)A)}$

```{r,echo=FALSE}
plot_96_profile(mut_mat[,c(1,7)])
```
In the blood-AC sample, A(C>T)A mutations account for approximately 0.05 of the total mutations.\
$Pr(SigA|A(C>T)A)=\frac{Pr(SigA)Pr(A(C>T)A|SigA)}{Pr(A(C>T)A)}=\frac{0.28(0.025)}{0.05}=\frac{0.007}{0.05}=0.14$

```{r,echo=FALSE}
plot_contribution(nmf_res$contribution, nmf_res$signature, mode = "absolute", coord_flip = TRUE)
```