### 2022-08-10 sandbox/June 25 Presentation Demo.Rmd

---
title: "Friday Presentation Demonstration"
author: "Audrey Mitchell, Marco Ruiz, Soua Yang"
date: "6/23/2021"
output: pdf_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(MutationalPatterns)
source("bayes_function.R")
source("simulation.R")
```

#Show Bayes function, explain inputs and outputs
function(mutation, sample, nmf_res)

#Testing our function by generating mutational profiles with varying proportions of distinct mutational signatures (Signature 18 and Signature 21) from the COSMIC library
```{r}
test_sample <- create_test_sample(c(18,21), c(  1,  0,
                                              0.9,0.1,
                                              0.7,0.3,
                                              0.5,0.5,
                                              0.3,0.7,
                                              0.1,0.9,
                                                0,  1),
                                  num = 7)
test_sample
```

```{r}
#Run NMF analysis on signatures
test_nmf <- extract_signatures(test_sample, rank = 2, nrun = 10)

#Plot signatures to match with signatures 18 and 21: V1 appears to match signature 21, V2 appears to match signature 18
plot_96_profile(test_nmf$signatures)

#set column and row names to names of signatures
colnames(test_nmf$signatures) <- c("Signature 18", "Signature 21")
rownames(test_nmf$contribution) <- c("Signature 18", "Signature 21")

plot_96_profile(test_nmf$signatures)
```

```{r}
test_nmf$contribution
plot_contribution(test_nmf$contribution, test_nmf$signature, mode = "relative", coord_flip = TRUE)
```

## Proportions should lean in favor for signature 21 and decrease with its sample size

```{r}
#G[T>C]A mutation is the most common in Signature 21
#269 times more likely in signature 21 than 18

sample.number <- c(1, 2, 3, 4, 5, 6, 7)
percent.sig21 <- c(0,10,30,50,70,90,100)
sig18.prob <- vector(mode = "double")
sig21.prob <- vector(mode = "double")

for (i in 1:7){
  probs <- extract_all_prob("G[T>C]A", i, test_nmf)
  sig18.prob[[i]]<- probs[1][[1]]
  sig21.prob[[i]]<- probs[2][[1]]
}

x <- data.frame(sample.number, percent.sig21, sig18.prob, sig21.prob)
colnames(x) <- c("Sample", "Percent Signature 21", "Signature 18 Posterior Probability", "Signature 21 Posterior Probability")
view(x)
```

## Proportions should lean in favor for signature 18 and decrease with its sample size

```{r}
sample.number2 <- c(1, 2, 3, 4, 5, 6, 7)
percent.sig21.2 <- c(0,10,30,50,70,90,100)
sig18.prob2 <- vector(mode = "double")
sig21.prob2 <- vector(mode = "double")

for (i in 1:7){
  probs <- extract_all_prob("G[C>A]A", i, test_nmf)
  sig18.prob2[[i]]<- probs[1][[1]]
  sig21.prob2[[i]]<- probs[2][[1]]
}

y <- data.frame(sample.number2, percent.sig21.2, sig18.prob2, sig21.prob2)
colnames(y) <- c("Sample", "Percent Signature 21", "Signature 18 Posterior Probability", "Signature 21 Posterior Probability")
view(y)
```

## Probabilities should almost reflect contribution proportions

```{r}
#low levels of C[C>T]G mutations in both signatures
#Sample 3 is 90% Signature 18 and 10% Signature 21
extract_all_prob("T[C>T]G", 2, test_nmf)
#Sample 4 is 90% Signature 21 and 10% Signature 18
extract_all_prob("T[C>T]G", 6, test_nmf)
```

#Example applying our function to real data
```{r}
# Loads a mutational profile with columns sample, ext.context and num
# Subsets such file to samples in sample.list and generates a matrix
# with the mutational profile that can be used by MutationalPatterns
load_mutational_matrix<- function(filename, sample.list) {

  mutational.signature <-  read_tsv(filename)
  order.tri <-  unique(mutational.signature$ext.context)
  sample.names <-  unique(mutational.signature$sample)

  mutational.signature <-  mutational.signature %>%
    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
    filter(sample %in% sample.list) %>%
    select(sample,ext.context,num) %>%
    spread(sample,num) %>%
    select(one_of("ext.context",sample.names))

  mut.matrix <- as.matrix(sapply(mutational.signature[,2:ncol(mutational.signature)], as.numeric))  
  rownames(mut.matrix) <-  mutational.signature$ext.context

  # Add 1e-04 for use with NMF
  mut.matrix <-  mut.matrix + 1e-04

  return (mut.matrix)
}

# Loads the published COSMIC mutational signatures in a format that can be
# used by MutationalPatterns
load_cosmic_matrix<- function(mut.order="supplied_results/mut_sig.order.txt") {

  sp_url <- paste("https://cancer.sanger.ac.uk/cancergenome/assets/",
                "signatures_probabilities.txt", sep = "")

  cancer_signatures <-  read.table(sp_url, sep = "\t", header = TRUE)
  # Match the order of the mutation types to MutationalPatterns standard
  mut_mat <-  read_tsv(mut.order)
  new_order <-  match(mut_mat$ext.context, cancer_signatures$Somatic.Mutation.Type) 
  # Reorder cancer signatures dataframe
  cancer_signatures <-  cancer_signatures[as.vector(new_order),]
  # Add trinucletiode changes names as row.names
  row.names(cancer_signatures) <-  cancer_signatures$Somatic.Mutation.Type
  # Keep only 96 contributions of the signatures in matrix
  cancer_signatures <-  as.matrix(cancer_signatures[,4:33])
  return(cancer_signatures)
}

cosmic.sigs <- load_cosmic_matrix()

mapping.stats <- read_tsv("supplied_results/master_statistics_table.txt") %>%
  select(sample.name,tumor.status,stat.Total_Reads)

samples.pilot <-  c("MSI_MLH1p_1")
samples.na.tissue <-  c("MSI_MLH1G_2","MSI_MLH1G_6","MSI_MSH2_11","MSS_1")
samples.exclude.age <-  c("MSI_MLH1HM_14")
samples.wrong.phenotype <-  c("MSI_MLH1G_9")
samples.small.bowel <- c("MSI_PMS2_12","MSS_9")

min.reads <- 1e7 #minimum number of reads

mapping.info.filter <-  mapping.stats %>%
  spread(tumor.status,stat.Total_Reads) %>%
  filter(!is.na(normal)) %>%
  filter(normal>min.reads & tumor>min.reads) %>%
  filter(!sample.name %in% samples.pilot) %>%
  filter(!sample.name %in% samples.na.tissue) %>%
  filter(!sample.name %in% samples.exclude.age) %>%
  filter(!sample.name %in% samples.wrong.phenotype) %>%
  filter(!sample.name %in% samples.small.bowel) 	
```

```{r}
mut.matrix <-  load_mutational_matrix("supplied_results/signatures.txt",
  mapping.info.filter$sample.name)
```

```{r}
nmf.res <- extract_signatures(mut.matrix, rank = 4, nrun = 10)
sig.names <-  c("PolE","Aging","FFPE","MMR")
colnames(nmf.res$signatures) <- sig.names
rownames(nmf.res$contribution) <- sig.names
```

```{r}
plot_96_profile(nmf.res$signatures)
```

```{r}
#plot_contribution(nmf.res$contribution, nmf.res$signature, mode = "relative", coord_flip = TRUE)
```

```{r}
#T[C>A]T is one of the peaks of the POLE mutational signature, MSS_6 is a sample with a high proportion of POLE mutations
extract_all_prob("T[C>A]T", "MSS_6", nmf.res)
```

```{r}
#G[C>T]T is a mutation that is present in all 4 mutational signatures, MSI_MLH1HM_15 is a sample with contributions from all 4
extract_all_prob("G[C>T]T", "MSI_MLH1HM_15", nmf.res)
```

#Simulating data for which the process behind each mutation is known

```{r}
#function to load ffpe signature
load_old_mutational_matrix<- function(filename,mut.order="supplied_results/mut_sig.order.txt") {

  mutational.signature = read_tsv(filename)
  mut_mat = read_tsv(mut.order)

  mut.join = mutational.signature %>%
    inner_join(mut_mat)

  order.tri = unique(mut.join$ext.context)

  mut.join = mut.join %>%
    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
    select(ext.context,prob)

  mut.matrix=as.matrix(sapply(mut.join[,2:ncol(mut.join)], as.numeric))  
  rownames(mut.matrix) = mut.join$ext.context

  return (t(mut.matrix)[,1:96])
}

#load ffpe signature
ffpe.signature <-  load_old_mutational_matrix("supplied_results/ffpe.signature.txt")
ffpe.signature.matrix <- as.matrix(ffpe.signature)
```

```{r}
#create ffpe sample vector of 1000 mutations
ffpe.sample.vector <- create_signature_sample_vector(ffpe.signature.matrix, 100)
ffpe.sample.vector
signature_cosine_similarity(ffpe.sample.vector, ffpe.signature.matrix)
```

```{r}
#create sample matrix to match ffpe signature
ffpe.sample.matrix <- create_signature_sample_matrix(ffpe.signature.matrix, 100)
ffpe.sample.matrix
#calculate the cosine similarity between sample matrix and original ffpe matrix
cos_sim_matrix(ffpe.sample.matrix, ffpe.signature.matrix)
#plot original ffpe matrix
plot_96_profile(ffpe.signature.matrix)
#plot our ffpe sample matrix
plot_96_profile(ffpe.sample.matrix)
```

```{r}
cosmic.sig6 <- as.matrix(cosmic.sigs[, 6])
sample.sig6 <- create_signature_sample_vector(cosmic.sig6, 200)

ffpe.20mutations <- create_signature_sample_vector(ffpe.signature.matrix, 20)

classification.df <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"), list(cosmic.sig6, ffpe.sig))

#view(classification.df)
mean(classification.df$misclassification)
```